### 2019-12-19 paper/paper.tex
Add BNF for model

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

This is a guide for authors who are preparing papers for JuliaCon using the \LaTeX{} document
preparation system and the \verb juliacon  class file.

\end{abstract}

\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

% using the backnaur package
\begin{bnf*}
    \bnfprod{model}{\bnfts{@model} \bnfsp \bnfpn{args} \bnfsp \bnfts{begin} \bnfsp \bnfpn{statements} \bnfsp \bnfpn{retn} \bnfsp \bnfts{end}}\\
    \bnfprod{args}{\bnfes \bnfor \bnfpn{Symbol} \bnfor \bnfpn{Symbol} \bnfts{,} \bnfpn{args}} \\
    \bnfprod{statements}{\bnfpn{statement} \bnfor \bnfpn{statement} \bnfts{\textbackslash n} \bnfpn{statements}} \\ 
    \bnfprod{statement}{\bnfpn{assign} \bnfor \bnfpn{sample}} \\
    \bnfprod{assign}{\bnfpn{Symbol} \bnfsp \bnfts{=} \bnfsp \bnfpn{Expr}} \\
    \bnfprod{sample}{\bnfpn{Symbol} \bnfsp \bnfts{\textasciitilde}} \bnfsp \bnfpn{Measure} \\
    \bnfprod{retn}{\bnfts{return} \bnfsp \bnfpn{Expr}} \\
    \bnfprod{Symbol}{\bnftd{Julia Symbol}} \\ 
    \bnfprod{Expr}{\bnftd{Julia Expr}} \\
    \bnfprod{Measure}{\bnftd{Probability measure (see text)}}
\end{bnf*}

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb \ref ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb juliacon.cls  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb juliacon  document class is implemented
as a complete class, not a document style option. In order to
use the \verb juliacon  document class, replace \verb article  by \verb juliacon  in the
\verb \documentclass  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb \documentclass{article}  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb \documentclass{juliacon}  \vskip 6pt
In general, the following standard document \verb style  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb 10pt,  \verb 11pt,  \verb 12pt   ? unavailable;
\item[(2)] \verb twoside  (no associated style file) ? \verb twoside  is the default;
\item[(3)] \verb fleqn, \verb leqno, \verb titlepage ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb juliacon  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb \begin{document} . For example, user-defined
\verb \def  and \verb \newcommand   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb \juliacon.cls  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb juliacon  class file includes the following features:
In general, once you have used the additional \verb juliacon.cls facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb \maketitle  command. For genaral tem and keywords use
\verb \terms ,
\verb \keywords  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb juliacon  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb juliacon  class file:
\begin{itemize}
\item \verb \section   \item \verb \subsection  \item \verb \subsubsection  \item \verb \paragraph  \end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb juliacon   class file provides unnumbered lists using the
\verb unnumlist  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb juliacon   class file also provides hyphen list using the
\verb itemize  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb juliacon   class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb figure   environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb \caption  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb \includegraphics   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb \label  and \verb \ref   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb Figure~\ref{sample-figure}   \subsection{Tables}
\label{subsub:sec_Tab}
The \verb juliacon   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb juliacon   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb rotating   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb juliacon class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb juliacon   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2019-12-21 paper/paper.tex
big updates

@@
-% JuliaCon proceedings template
-\documentclass{juliacon}
-\setcounter{page}{1}
+%%%% Small single column format
+\documentclass[anonymous=false, %
+               format=acmsmall, %
+               review=true, %
+               screen=true, %
+               nonacm=true]{acmart}
+
+\usepackage[ruled]{algorithm2e} 
+%\usepackage{parskip}
+\usepackage{backnaur}
+
+\urlstyle{tt}
+\citestyle{acmauthoryear}
 
@@
 
-\input{header}
+\title{Soss}
+%  \titlenote{This is a titlenote}
+%  \subtitle{This is a subtitle}
+%  \subtitlenote{Subtitle note}
+
+\author{Chad Scherrer}
+\orcid{0000-0002-1490-0304}
+\affiliation{%
+  \institution{RelationalAI}
+  % \department{Department of Brain and Cognitive Sciences}
+  %\streetaddress{43 Vassar St}
+  %\city{Cambridge}
+  %\state{MA}
+  %\postcode{02139}
+  %\country{USA}
+}
+\email{chad.scherrer@gmail.com}
+
+\author{Taine Zhao}
+%\orcid{1234-5678-9012-3456}
+\affiliation{%
+  \institution{University of Tsukuba}
+  \department{Department of Computer Science}
+  %\streetaddress{625 Mt Auburn St #3}
+  %\city{Cambridge}
+  %\state{MA}
+  %\postcode{02138}
+  %\country{USA}
+}
+% \email{apfeffer@cra.com}
+%\renewcommand\shortauthors{Mage, M. et al}
+
+\begin{abstract}
+Abstract goes here
+\end{abstract}
 
@@
 
-\begin{abstract}
-
-This is a guide for authors who are preparing papers for JuliaCon using the \LaTeX{} document
-preparation system and the \verb juliacon  class file.
-
-\end{abstract}
-
 \section{Introduction}
 
-The \LaTeX{} document preparation system is a special version of the
-\TeX; typesetting program where-in a collection of \TeX; commands
-are added to \LaTeX{} to simplify typesetting. Importantly, it allows
-the author to concentrate on the logical structure of the document
-rather than its visual layout.\vskip 6pt
-Moreover, \LaTeX{} provides a consistent and comprehensive document
-preparation interface. There are simple-to-use commands for
-generating a table of contents, lists of figures and/or tables, and indexes.
-\LaTeX{} can automatically number list entries, equations, figures,
-tables, and footnotes, as well as articles, sections, and subsections.
-Using this numbering system, bibliographic citations, page
-references, and cross references to any other numbered entity (e.g.
-article, section, equation, figure, list entry, etc.) become quite simple
-and straightforward. The use of \LaTeX{} document classes allows
-a simple change of class to transform the appearance of your document.\vskip 6pt
-\LaTeX{} is a powerful tool for managing long and complex documents.
-In particular, partial processing enables long documents to
-be produced article by article without losing sequential information.
-The use of document classes allows a simple change of style
-(or style option) to transform the appearance of your document.
+\section{End Matter}
 
+Identification of funding sources and other support, and thanks to
+individuals and groups that assisted in the research and the
+preparation of the work should be included in an acknowledgment
+section, which is placed just before the reference section in your
+document. This section has a special environment:
+\begin{verbatim}
+  \begin{acks}
+  ...
+  \end{acks}
+\end{verbatim}
+This ensure that information contained therein can be more easily collected during the article metadata extraction phase, and ensures consistency in the spelling of the section heading.
+An example is below.
+
+The \verb|\appendix| command is used to start the appendix sections in the document, which are indexed with letters rather than numbers.
+This command should be included after  the \verb|\bibliography| command, which creates the references section.
+
+\section{Model syntax}
+
+\begin{figure}[!t]
+  \centering
+  %\fbox{\rule[-.5cm]{0cm}{\linedwid} \rule[-.5cm]{4cm}{0cm}}
   % using the backnaur package
@@
 \end{bnf*}
-
-\section{The JuliaCon Article Class}
-\label{sec:documentclass}
-%
-The juliacon class file preserves the standard LATEX{} interface such
-that any document that can be produced using the standard LATEX{}
-article class can also be produced with the class file.\vskip 6pt
-It is likely that the make up will change after file submission. For
-this reason, we ask you to ignore details such as slightly long lines,
-page stretching, or figures falling out of synchronization, as these
-details can be dealt with at a later stage.\vskip 6pt
-Use should be made of symbolic references (\verb \ref ) in order to
-protect against late changes of order, etc.
-
-\section{USING THE JuliaCon Article CLASS FILE}
-
-If the file \verb juliacon.cls  is not already in the appropriate system directory
-for \LaTeX{} files, either arrange for it to be put there or copy
-it to your working directory. The \verb juliacon  document class is implemented
-as a complete class, not a document style option. In order to
-use the \verb juliacon  document class, replace \verb article  by \verb juliacon  in the
-\verb \documentclass  command at the beginning of your document:
-\vskip 6pt
-\begin{centering}
-    \verb \documentclass{article}  \end{centering}
-\vskip 6pt
-replace by
-\vskip 6pt
- \verb \documentclass{juliacon}  \vskip 6pt
-In general, the following standard document \verb style  options should
-{ \itshape not} be used with the {\footnotesize \itshape article} class file:
-\begin{enumerate}
-\item[(1)] \verb 10pt,  \verb 11pt,  \verb 12pt   ? unavailable;
-\item[(2)] \verb twoside  (no associated style file) ? \verb twoside  is the default;
-\item[(3)] \verb fleqn, \verb leqno, \verb titlepage ? should not be used;
-\end{enumerate}
-
-\section{Additional Document Style Options}
-\label{sec:additional_doc}
-%
-The following additional style option is available with the \verb juliacon  class file:
-\vskip 6pt
-Please place any additional command definitions at the very start of
-the \LaTeX{} file, before the \verb \begin{document} . For example, user-defined
-\verb \def  and \verb \newcommand   commands that define macros for
-technical expressions should be placed here. Other author-defined
-macros should be kept to a minimum.
-\vskip 6pt
-Commands that differ from the standard \LaTeX{} interface, or that
-are provided in addition to the standard interface, are explained in
-this guide. This guide is not a substitute for the \LaTeX{} manual itself.
-Authors planning to submit their papers in \LaTeX{} are advised to use
-\verb \juliacon.cls  as early as possible in the creation of their files.
-
-%
-%
-%
-%
-\begin{table*}[t]
-\tabcolsep22pt
-\tbl{If necessary, the tables can be extended both columns.}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\label{tab:symbols}
-\begin{tabnote}
-This is an example of table footnote.
-\end{tabnote}
-\end{table*}
-% \begin{figure*}[t]
-% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
-% \caption{If necessary, the images can be extended both columns.}
-%   \label{fig:sample_image}
-% \end{figure*}
-
-\section{Additional features}
-\label{sec:additional_faci}
-In addition to all the standard \LaTeX{} design elements, the \verb juliacon  class file includes the following features:
-In general, once you have used the additional \verb juliacon.cls facilities
-in your document, do not process it with a standard \LaTeX{} class
-file.
-
-\subsection{Titles, Author's Name, and Affiliation}
-\label{subsub:title_auth}
-The title of the article, author's name, and affiliation are used at the
-beginning of the article (for the main title). These can be produced
-using the following code:
-
-\begin{verbatim}
-\title{ This is an example of article title} }
-\author{
-   \large 1st Author \\[-3pt]
-   \normalsize 1st author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	1st author's email address \\[-3pt]
-  \and
-   \large 2nd Author \\[-3pt]
-   \normalsize 2nd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	2nd author's email address \\[-3pt]
-\and
-   \large 3rd Author \\[-3pt]
-   \normalsize 3rd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	3rd author's email address \\[-3pt]
-}
-\maketitle
-\end{verbatim}
-
-\subsection{Writing Julia code}
-
-A special environment is already defined for Julia code,
-built on top of \textit{listings} and \textit{jlcode}.
-
-\begin{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-\end{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-
-\subsection{Abstracts, Key words, term etc...}
-\label{subsub:abs_key_etc}
-
-At the beginning of your article, the title should be generated
-in the usual way using the \verb \maketitle  command. For genaral tem and keywords use
-\verb \terms ,
-\verb \keywords  commands respectively. The abstract should be enclosed
-within an abstract environment, All these environment
-can be produced using the following code:
-\begin{verbatim}
-\terms{Experimentation, Human Factors}
-
-\keywords{Face animation, image-based modelling...}
-
-\begin{abstract}
-In this paper, we propose a new method for the
-systematic determination of the model's base of
-time varying delay system. This method based on
-the construction of the classification data related
-to the considered system. The number, the orders,
-the time delay and the parameters of the local
-models are generated automatically without any
-knowledge about the full operating range of the
-process. The parametric identification of the local
-models is realized by a new recursive algorithm for
-on line identification of systems with unknown time
-delay. The proposed algorithm allows simultaneous
-estimation of time delay and parameters of
-discrete-time systems. The effectiveness of
-the new method has been illustrated through
-simulation.
-\end{abstract}
-
-\end{verbatim}
-
-\section{Some guidelines}
-\label{sec:some_guide}
-The following notes may help you achieve the best effects with the
-\verb juliacon  class file.
-
-\subsection{Sections}
-\label{subsub:sections}
-\LaTeXe{}  provides four levels of section headings and they are all
-defined in the \verb juliacon  class file:
-\begin{itemize}
-\item \verb \section   \item \verb \subsection  \item \verb \subsubsection  \item \verb \paragraph  \end{itemize}
-Section headings are automatically converted to allcaps style.
-\subsection{Lists}
-\label{sec:lists}
-%
-The \verb juliacon   class file provides unnumbered lists using the
-\verb unnumlist  environment for example,
-
-\begin{unnumlist}
-\item First unnumbered item which has no label and is indented from the
-left margin.
-\item Second unnumbered item.
-\item Third unnumbered item.
-\end{unnumlist}
-The unnumbered list which has no label and is indented from the
-left margin. was produced by:
-\begin{verbatim}
-\begin{unnumlist}
-\item First unnumbered item...
-\item Second unnumbered item...
-\item Third unnumbered item...
-\end{unnumlist}
-\end{verbatim}
-
-The \verb juliacon   class file also provides hyphen list using the
-\verb itemize  environment for example,
-\begin{itemize}
-\item First unnumbered bulleted item which has no label and is indented
-from the left margin.
-\item Second unnumbered bulleted item.
-\item Third unnumbered bulleted item which has no label and is indented
-from the left margin.
-\end{itemize}
-was produced by:
-\begin{verbatim}
-\begin{itemize}
-\item First item...
-\item Second item...
-\item Third item...
-\end{itemize}
-\end{verbatim}
-
-Numbered list is also provided in acmtog class file using the
-enumerate environment for example,
-\begin{enumerate}
-\item The attenuated and diluted stellar radiation.
-\item Scattered radiation, and
-\item Reradiation from other grains.
-\end{enumerate}
-
-was produced by:
-\begin{verbatim}
-\begin{enumerate}
-\item The attenuated...
-\item Scattered radiation, and...
-\item Reradiation from other grains...
-\end{enumerate}
-\end{verbatim}
-\subsection{Illustrations (or figures)}
-\label{subsub:sec_Illus}
-The \verb juliacon   class file will cope with most of the positioning of
-your illustrations and you should not normally use the optional positional
-qualifiers on the \verb figure   environment that would override
-these decisions.
-\vskip 6pt
-
-%
-\begin{figure}[t]
-\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
-\caption{This is example of the image in a column.}
-	\label{fig:sample_figure}
+  \caption{Backus-Naur Form representation for a user-specified model.}
+  \label{fig:bnf}
+  \Description{Placeholder figure.}
 \end{figure}
 
-The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
-organization \footnote{https://github.com/JuliaGraphs}.
+\section{Implementation}
 
-Figure captions should be \emph{below} the figure itself, therefore the
-\verb \caption  command should appear after the figure or space left for
-an illustration. For example, Figure 1 is produced using the following
-commands:
+\section{Performance}
 
-\begin{verbatim}
-\begin{figure}
-\centerline{\includegraphics[width=20pc]{Graphics.eps}}
-\caption{An example of the testing process for a
-binary tree. The globa null hypothesis is tested
-first at level $\alpha$ (a), and the level of
-individual variables is reached last (d). Note
-that individual hypotheses can be tested at
-level $\alpha/4$ and not $\alpha/8$ as one might
-expect at first.}
-\label{sample-figure_2}
-\end{figure}
-\end{verbatim}
-Figures can be resized using first and second argument of
-\verb \includegraphics   command. First argument is used for modifying
-figure height and the second argument is used for modifying
-figure width respectively.
-\vskip 6pt
-Cross-referencing of figures, tables, and numbered, displayed
-equations using the \verb \label  and \verb \ref   commands is encouraged.
-For example, in referencing Figure 1 above, we used
-\verb Figure~\ref{sample-figure}   \subsection{Tables}
-\label{subsub:sec_Tab}
-The \verb juliacon   class file will cope with most of the positioning of
-your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
-decisions. Table captions should be at the top.
-\begin{verbatim}
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Train70 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test70 & Testing Data I & & 105\\\hline
-Test30 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}\end{verbatim}
+\begin{acks}
+We would like to acknowledge...
+\end{acks}
 
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\subsection{Landscaping Pages}
-\label{subsub:landscaping_pages}
-If a table is too wide to fit the standard measure, it may be turned,
-with its caption, to 90 degrees. Landscape tables cannot be produced
-directly using the \verb juliacon   class file because \TeX{} itself cannot
-turn the page, and not all device drivers provide such a facility.
-The following procedure can be used to produce such pages.
-\vskip 6pt
-Use the package \verb rotating   in your document and change the coding
-from
-\begin{verbatim}
-\begin{table}...\end{table}
-to
-\begin{sidewaystable}...\end{sidewaystable}
-and for figures
-\begin{figure}...\end{figure}
-to
-\begin{sidewaysfigure}...\end{sidewaysfigure}
-\end{verbatim}
+\bibliographystyle{acm-reference-format}
+\bibliography{probprog-2020-instructions}
 
-environments in your document to turn your table on the appropriate
-page of your document. For instance, the following code prints
-a page with the running head, a message half way down and the
-table number towards the bottom.
-\begin{verbatim}
-\begin{sidewaystable}
-\tbl{Landscape table caption to go here.}{...}
-\label{landtab}
-\end{sidewaystable}
-\end{verbatim}
+\appendix
 
-\subsection{Double Column Figure and Tables}
-\label{subsub:double_fig_tab}
-For generating the output of figures and tables in double column
-we can use the following coding:
+\section{Appendices}
 
-\begin{enumerate}
-\item For Figures:
-\begin{verbatim}
-\begin{figure*}...\end{figure*}
-\end{verbatim}
-\item For landscape figures:
-\begin{verbatim}
-\begin{sidewaysfigure*}...\end{sidewaysfigure*}
-\end{verbatim}
-\item For Tables:
-\begin{verbatim}
-\begin{table*}...\end{table*}
-\end{verbatim}
-\item For landscape tables:
-\begin{verbatim}
-\begin{sidewaystable*}...\end{sidewaystable*}
-\end{verbatim}
-\end{enumerate}
-
-\subsection{Typesetting Mathematics}
-\label{subsub:type_math}
-The \verb juliacon class file will set displayed mathematics with center to
-the column width, provided that you use the \LaTeXe{} standard of
-open and closed square brackets as delimiters.
-The equation
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-
-was typeset using the acmtog class file with the commands
-
-\begin{verbatim}
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-\end{verbatim}
-
-For display equations, cross-referencing is encouraged. For example,
-\begin{verbatim}
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-\end{verbatim}
-The following output is generated with the above coding:
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-
-\subsection{Enunciations}
-\label{subsub:enunciation}
-The \verb juliacon   class file generates the enunciations with the help of
-the following commands:
-\begin{verbatim}
-\begin{theorem}...\end{theorem}
-\begin{strategy}...\end{strategy}
-\begin{property}...\end{property}
-\begin{proposition}...\end{proposition}
-\begin{lemma}...\end{lemma}
-\begin{example}...\end{example}
-\begin{proof}...\end{proof}
-\begin{definition}...\end{definition}
-\begin{algorithm}...\end{algorithm}
-\begin{remark}...\end{remark}
-\end{verbatim}
-The above-mentioned coding can also include optional arguments
-such as
-\begin{verbatim}
-\begin{theorem}[...]. Example for theorem:
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven ... created equal.
-\end{theorem}
-\end{verbatim}
-
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven years ago our fathers brought forth,
-upon this continent, a new nation, conceived in Liberty,
- and dedicated to the proposition that all men are
-created equal.
-\end{theorem}
-
-\subsection{Extract}
-\label{subsub:extract}
-Extract environment should be coded within
-\begin{verbatim}
-\begin{extract}..\end{extract}
-\end{verbatim}
-
-\subsection{Balancing column at last page}
-\label{subsub:Balance}
-For balancing the both column length at last page use :
-\begin{verbatim}
-\vadjust{\vfill\pagebreak}
-\end{verbatim}
-
-%\vadjust{\vfill\pagebreak}
-
-at appropriate place in your \TeX{} file or in bibliography file.
-
-\section{Handling references}
-\label{subsub:references}
-References are most easily (and correctly) generated using the
-BIBTEX, which is easily invoked via
-\begin{verbatim}
-\bibliographystyle{juliacon}
-\bibliography{ref}
-\end{verbatim}
-When submitting the document source (.tex) file to external
-parties, the ref.bib file should be sent with it.
-\cite{bezanson2017julia}
-
-\input{bib.tex}
-
+Authors may provide appendices to accompany their extended abstract submissions. Please note however that reviewers will not be required to comment on material in these appendices.
 \end{document}
-
-% Inspired by the International Journal of Computer Applications template

### 2019-12-21 style/probprog-2020-instructions.tex
big updates

%%%% Small single column format
\documentclass[anonymous=false, %
               format=acmsmall, %
               review=true, %
               screen=true, %
               nonacm=true]{acmart}

\usepackage[ruled]{algorithm2e} 
%\usepackage{parskip}

\urlstyle{tt}
\citestyle{acmauthoryear}

\begin{document}

\title{Formatting Instructions for PROBPROG 2020 Abstracts}
%  \titlenote{This is a titlenote}
%  \subtitle{This is a subtitle}
%  \subtitlenote{Subtitle note}

\author{Vikash Mansinghka}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{Massachussets Institute of Technology}
  \department{Department of Brain and Cognitive Sciences}
  %\streetaddress{43 Vassar St}
  %\city{Cambridge}
  %\state{MA}
  %\postcode{02139}
  %\country{USA}
}
\email{vkm@mit.edu}

\author{Avi Pfeffer}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{Charles River Analytics}
  %\streetaddress{625 Mt Auburn St #3}
  %\city{Cambridge}
  %\state{MA}
  %\postcode{02138}
  %\country{USA}
}
\email{apfeffer@cra.com}

\author{Jean-Baptiste Tristan}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{Oracle Labs}
  %\streetaddress{15 Network Dr}
  %\city{Burlington}
  %\state{MA}
  %\postcode{01803}
  %\country{USA}
}
\email{jean.baptiste.tristan@oracle.com}

\author{Jan-Willem van de Meent}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{Northeastern University}
  \department{Khoury College of Computer Sciences}
  %\streetaddress{360 Huntington Ave}
  \city{Boston}
  \state{MA}
  %\postcode{02115}
  \country{USA}
}
\email{j.vandemeent@northeastern.edu}
%\renewcommand\shortauthors{Mage, M. et al}

\begin{abstract}
These are formatting instructions for extended abstract submissions to the International Conference on Probabilistic Programmming (PROBPROG) 2020. Extended abstracts, which are non-archival, will be reviewed for consideration for a poster presentation or an oral presentation at the conference. Following the conference, authors of a selected abstracts will be invited to submit full-length papers to \emph{Probabilistic Programming}, a new online journal that accompanies the conference.
\end{abstract}

\maketitle

\section{Length and Style}

Extended abstracts may be 2--6 pages in length, exclusive of references. There is no length limitation on references. Authors may include supplementary material, which may be uploaded as part of a single PDF, but reviewers will not be required to comment on this material.

Authors are encouraged to make use of these style files, which are available at:
\begin{center}
   \url{https://probprog.cc/2020/probprog-2020-style.zip}
\end{center}
These files consist of:
\begin{itemize}
\item[-] \verb+probprog-2020-instructions.pdf+, which  contains these instructions and illustrates the format for PROBPROG extended abstracts.

\item[-] \verb+probprog-2020-instructions.tex+, which may be used as a template for your submission.

\item[-] \verb+probprog-2020-instructions.bib+, which provides sample references.

\item[-] \verb+acmart.cls+. The LaTeX class developed by the Association for Computing Machinery (ACM).

\item[-] \verb+acm-reference-format.{bbx,bst,cbx,dbx}+. The default reference format. 
\end{itemize}

\section{Submission}

Abstracts should be submitted via the PROBPROG 2020 CMT site:
\begin{center}
    \url{https://cmt3.research.microsoft.com/PROBPROG2020/}
\end{center}
Please ensure that the list and ordering of authors in the CMT record matches that of the submitted PDF file. This record may be used for purposes such as creating a listing on the conference website.

\section{Formatting}

\subsection{Title, Author List, and Abstract}

This year, submissions do not have to be anonymous. Authors may optionally use the \verb|abstract| environment to include an overview of the submission. Author postal and e-mail addresses are listed at the bottom of the first page.

\subsection{Headings}

Sections should be numbered and use standard LaTex sectioning commands:
\verb|\section|, \verb|\subsection|, \verb|\subsubsection|, and
\verb|\paragraph|.

\subsection{Citations}

Authors are allowed to use a citation style of their choosing. The use of BibTex for is strongly recommended. We provide the default ACM reference format as part of these style files. To  list references according to this format, include the following two lines just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{acm-reference-format}
  \bibliography{bibfile}
\end{verbatim}
Here ``\verb|bibfile|'' is the name, without the ``\verb|.bib|'' suffix, of the BibTex file.

Citations and references are in ``author year'' style by default. Citations can either be listed in parentheses, e.g.~\cite{probprog2020instructions}, by using the standard \verb|\cite| command, or inline in the text, e.g.~\citet{probprog2020instructions}, by using the \verb|\citet| command.

\subsection{Tables}

The ``\verb|acmart|'' style includes the ``\verb|booktabs|''
package \cite{Fear05} for preparing
high-quality tables. Table captions are placed {\itshape above} the table. See Table~\ref{tab:freq} for an example.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \verb|table| to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\verb|tabular| environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \verb|tabular| material are found in the
\textit{LaTeX\ User's Guide}.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \verb|table*| to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \emph{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \verb|equation|
environment. An unnumbered display equation is produced by the
\emph{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \emph{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate LaTeX's able handling of numbering.

\subsection{Figures}

The \verb|figure| environment should be used for figures. One or more images can be placed within a figure. Your figures should contain a caption which describes the figure to the reader. Figure captions are placed {\itshape below} the figure. See Figure~\ref{fig:example} for an example.

\begin{figure}[!t]
  \centering
  %\fbox{\rule[-.5cm]{0cm}{\linedwid} \rule[-.5cm]{4cm}{0cm}}
  \framebox[0.8\linewidth]{\rule{0pt}{1.5in}}
  \caption{Figure captions are positioned below the image.}
  \label{fig:example}
  \Description{Placeholder figure.}
\end{figure}

% \subsection{Citations within the text}

% Citations within the text should be based on the \texttt{natbib} package
% and include the authors' last names and year (with the ``et~al.'' construct
% for more than two authors). When the authors or the publication are
% included in the sentence, the citation should not be in parenthesis (as
% in ``See \citet{Hinton06} for more information.''). Otherwise, the citation
% should be in parenthesis (as in ``Deep learning shows promise to make progress towards AI~\citep{Bengio+chapter2007}.'').

% The corresponding references are to be listed in alphabetical order of
% authors, in the \textsc{References} section. As to the format of the
% references themselves, any style is acceptable as long as it is used
% consistently.

% \subsection{Footnotes}

% Indicate footnotes with a number\footnote{Sample of the first footnote} in the
% text. Place the footnotes at the bottom of the page on which they appear.
% Precede the footnote with a horizontal rule of 2~inches
% (12~picas).\footnote{Sample of the second footnote}

% \subsection{Figures}

% All artwork must be neat, clean, and legible. Lines should be dark
% enough for purposes of reproduction; art work should not be
% hand-drawn. The figure number and caption always appear after the
% figure. Place one line space before the figure caption, and one line
% space after the figure. The figure caption is lower case (except for
% first word and proper nouns); figures are numbered consecutively.

% Make sure the figure caption does not get separated from the figure.
% Leave sufficient space to avoid splitting the figure and figure caption.

% You may use color figures.
% However, it is best for the
% figure captions and the paper body to make sense if the paper is printed
% either in black/white or in color.
% \begin{figure}[h]
% \begin{center}
% %\framebox[4.0in]{$\;$}
% \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
% \end{center}
% \caption{Sample figure caption.}
% \end{figure}

% \subsection{Tables}

% All tables must be centered, neat, clean and legible. Do not use hand-drawn
% tables. The table number and title always appear before the table. See
% Table~\ref{sample-table}.

% Place one line space before the table title, one line space after the table
% title, and one line space after the table. The table title must be lower case
% (except for first word and proper nouns); tables are numbered consecutively.

% \begin{table}[t]
% \caption{Sample table title}
% \label{sample-table}
% \begin{center}
% \begin{tabular}{ll}
% \toprule
% \multicolumn{1}{c}{\bf Part}  &\multicolumn{1}{c}{\bf Description}\\ 
% \midrule
% Dendrite         &Input terminal \\
% Axon             &Output terminal \\
% Soma             &Cell body (contains cell nucleus) \\
% \bottomrule
% \end{tabular}
% \end{center}
% \end{table}

\section{End Matter}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document. This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
This ensure that information contained therein can be more easily collected during the article metadata extraction phase, and ensures consistency in the spelling of the section heading. An example is below.

The \verb|\appendix| command is used to start the appendix sections in the document, which are indexed with letters rather than numbers. This command should be included after  the \verb|\bibliography| command, which creates the references section.

\begin{acks}
We would like to acknowledge Boris Veytsman and the Association for Computing Machinery for creating the excellent ACM article class that we use for PROBPROG submissions.
\end{acks}

\bibliographystyle{acm-reference-format}
\bibliography{probprog-2020-instructions}

\appendix

\section{Appendices}

Authors may provide appendices to accompany their extended abstract submissions. Please note however that reviewers will not be required to comment on material in these appendices.
\end{document}

### 2019-12-22 paper/paper.tex
code example
edits

@@
 \usepackage{backnaur}
+\usepackage{jlcode}
+\usepackage{todonotes}
 
@@
 
+Here's some example code. Lots of dead space on the right, maybe want to try to keep code in a multi-column figure
+
+\begin{lstlisting}[language = Julia]
+  using Plots
+
+  x = -3.0:0.01:3.0
+  y = rand(length(x))
+  plot(x, y)
+\end{lstlisting}
+
 \section{End Matter}
@@
 
+\begin{lstlisting}[language=julia]
+struct Model{A,B,M} 
+    args  :: Vector{Symbol}
+    vals  :: NamedTuple
+    dists :: NamedTuple
+    retn  :: Union{Nothing, Symbol, Expr}
+end
+\end{lstlisting}
+
 \section{Performance}
+
+\todo[inline]{Can Soss really generate arbitrary code?}
+The ability to generate arbitrary  Julia code makes it difficult to measure performance in Soss.
+When a performance bottleneck is found, developers or users can ask Soss to generate something different.
+This design places no a priori constraints on performance.
+
+\section{Extending Soss}
 

### 2019-12-22 paper/paper.tex
Start on related work

@@
 
+\section{Related Work}
+
+The idea of code generation and symbolic simplification in an embedded PPL goes back to \emph{Passage} \cite{Scherrer2012}.
+
+\emph{Hakaru} \cite{narayanan2016probabilistic} takes a more ambitious symbolic approach in a stand-alone PPL, allowing a wider variety of program transformations.
+
+Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspried by \emph{Stan} \cite{stan:2017}.
+
+\emph{Gen} \cite{Cusumano-Towner:2019}
+
+\emph{Turing} \cite{ge2018t}
+
 \begin{acks}
@@
 \bibliographystyle{acm-reference-format}
-\bibliography{probprog-2020-instructions}
+\bibliography{ref}
 

### 2019-12-22 paper/paper.tex
Just some writing

@@
 \begin{abstract}
-Abstract goes here
+Soss is a Julia-based embedded probabilistic programming language.
 \end{abstract}
@@
 
+\section{Model Transformations}
+
+\subsection{Markov Blankets}
+
+\subsection{Conditional Predictive models}
+
+\subsection{Do}
+
+\section{Symbolic Simplifcation}
+
+\section{Code Generation}
+
+\section{Implementation}
+
+\begin{lstlisting}[language=julia]
+struct Model{A,B,M} 
+    args  :: Vector{Symbol}
+    vals  :: NamedTuple
+    dists :: NamedTuple
+    retn  :: Union{Nothing, Symbol, Expr}
+end
+\end{lstlisting}
+
+\section{Performance}
+
+\todo[inline]{Let's implement models from \url{https://statisticalrethinkingjulia.github.io/MCMCBenchmarks.jl/latest/benchmarks/} for benchmark comparisons}
+
+\todo[inline]{Can Soss really generate arbitrary code?}
+
+The ability to generate arbitrary  Julia code makes it difficult to measure performance in Soss.
+When a performance bottleneck is found, developers or users can ask Soss to generate something different.
+This design places no a priori constraints on performance.
+
+\section{Extending Soss}
+
+\section{Related Work}
+
+The idea of code generation and symbolic simplification in an embedded PPL goes back to \emph{Passage} \cite{Scherrer2012}.
+
+\emph{Hakaru} \cite{narayanan2016probabilistic} takes a more ambitious symbolic approach in a stand-alone PPL, allowing a wider variety of program transformations.
+
+Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspried by \emph{Stan} \cite{stan:2017}.
+
+\emph{Gen} \cite{Cusumano-Towner:2019} was developed independently from Soss, but takes a similar approach (and distinct from most PPLs) in its representation of a model as a function from its arguments to a "trace" (to use Gen's terminology).
+In both Soss and Gen's static DSL, this trace is a mapping from variable names to values.
+The similarity of these systems makes interoperability relatively straightforward, as demonstrated in \emph{SossGen}(\url{https://github.com/cscherrer/SossGen.jl}).
+
+\emph{Turing} \cite{ge2018t}
+
+\begin{acks}
+We would like to acknowledge...
+\end{acks}
+
+\bibliographystyle{acm-reference-format}
+\bibliography{ref}
+
+\appendix
+
+\section{Model Syntax}
+
 \begin{figure}[!t]
@@
 
-\section{Implementation}
-
-\begin{lstlisting}[language=julia]
-struct Model{A,B,M} 
-    args  :: Vector{Symbol}
-    vals  :: NamedTuple
-    dists :: NamedTuple
-    retn  :: Union{Nothing, Symbol, Expr}
-end
-\end{lstlisting}
-
-\section{Performance}
-
-\todo[inline]{Can Soss really generate arbitrary code?}
-The ability to generate arbitrary  Julia code makes it difficult to measure performance in Soss.
-When a performance bottleneck is found, developers or users can ask Soss to generate something different.
-This design places no a priori constraints on performance.
-
-\section{Extending Soss}
-
-\section{Related Work}
-
-The idea of code generation and symbolic simplification in an embedded PPL goes back to \emph{Passage} \cite{Scherrer2012}.
-
-\emph{Hakaru} \cite{narayanan2016probabilistic} takes a more ambitious symbolic approach in a stand-alone PPL, allowing a wider variety of program transformations.
-
-Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspried by \emph{Stan} \cite{stan:2017}.
-
-\emph{Gen} \cite{Cusumano-Towner:2019}
-
-\emph{Turing} \cite{ge2018t}
-
-\begin{acks}
-We would like to acknowledge...
-\end{acks}
-
-\bibliographystyle{acm-reference-format}
-\bibliography{ref}
-
-\appendix
-
-\section{Appendices}
-
 Authors may provide appendices to accompany their extended abstract submissions. Please note however that reviewers will not be required to comment on material in these appendices.

### 2019-12-25 paper/paper.tex
abstract

@@
 \begin{abstract}
-Soss is a Julia-based embedded probabilistic programming language.
+Soss is a probabilistic programming language embedded probabilistic programming language embedded in the Julia language.
+Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without reqquiring casual users to worry about such details.
 \end{abstract}
@@
 \section{Introduction}
+
+There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most in a number of ways...
 
@@
 
-\section{End Matter}
-
-Identification of funding sources and other support, and thanks to
-individuals and groups that assisted in the research and the
-preparation of the work should be included in an acknowledgment
-section, which is placed just before the reference section in your
-document. This section has a special environment:
-\begin{verbatim}
-  \begin{acks}
-  ...
-  \end{acks}
-\end{verbatim}
-This ensure that information contained therein can be more easily collected during the article metadata extraction phase, and ensures consistency in the spelling of the section heading.
-An example is below.
-
-The \verb|\appendix| command is used to start the appendix sections in the document, which are indexed with letters rather than numbers.
-This command should be included after  the \verb|\bibliography| command, which creates the references section.
-
 \section{Model syntax}
@@
 
-\subsection{Do}
+\subsection{Causal Interventions (``Do'' operator)}
 
@@
 In both Soss and Gen's static DSL, this trace is a mapping from variable names to values.
-The similarity of these systems makes interoperability relatively straightforward, as demonstrated in \emph{SossGen}(\url{https://github.com/cscherrer/SossGen.jl}).
+The similarity of these systems makes interoperability relatively straightforward, as demonstrated in \emph{SossGen} (\url{https://github.com/cscherrer/SossGen.jl}).
 

### 2019-12-29 paper/paper.tex
edits

@@
 \begin{abstract}
-Soss is a probabilistic programming language embedded probabilistic programming language embedded in the Julia language.
+We present Soss, a probabilistic programming language embedded in the Julia language.
 Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without reqquiring casual users to worry about such details.
+
+The approach taken by Soss makes it easy to extend to take advantage of other packages in the rapidly-growing Julia ecosystem.
+At the time of this writing, Soss users can choose from several inference back-ends and connect easily with larger systems SymPy and Gen.
 \end{abstract}
@@
 \begin{lstlisting}[language = Julia]
-  using Plots
-
-  x = -3.0:0.01:3.0
-  y = rand(length(x))
-  plot(x, y)
+m = @model n begin
+    x ~ Normal(0,10) |> iid(n)
+    α ~ Normal()
+    β ~ Normal()
+    σ ~ HalfNormal()
+    yhat = α .+ β .* x
+    y ~ For(n) do j 
+            Normal(yhat[j],σ)
+        end
+end
 \end{lstlisting}
@@
 
-\section{Symbolic Simplifcation}
+\section{Symbolic Manipulation}
 
-\section{Code Generation}
+\subsection{Symbolic Log-density}
+
+\[
+  \ell =
+  - 0.5 α^{2} 
+  - 0.5 β^{2}
+  - σ^{2} 
+  + \sum_{\left(j_{1}\right)=1}^{n} \left(- \log{\left(\sigma \right)} 
+  - \frac{\log{\left(\pi \right)}}{2} 
+  - \frac{\log{\left(2 \right)}}{2} 
+  - \frac{\left({y}_{\left(j_{1}\right)} 
+  - {\hat{y}}_{\left(j_{1}\right)}\right)^{2}}{2 \sigma^{2}}\right) 
+  + \sum_{j_{1}=1}^{n} \left(- 0.005 {x}_{j_{1}}^{2} 
+  - 2.99573227355399 
+  - \frac{\log{\left(\pi \right)}}{2} 
+  + \frac{\log{\left(2 \right)}}{2}\right) 
+  - 2 \log{\left(\pi \right)} 
+  - 1.38629436111989
+\]
+
+\[
+  \ell =
+  - 3.7 
+  - 4.1 n 
+  - \frac{ \alpha^{2} }{2}
+  - \frac{\beta^2}{2}
+  - \sigma^{2} 
+  - n \log{\left(\sigma \right)} 
+  - 0.005 \sum_{j_{1}=1}^{n} {x}_{j_{1}}^{2} 
+  - \frac{1}{\sigma^2}\sum_{j_{1}=1}^{n} \left({y}_{j_{1}} 
+  - {\hat{y}}_{j_{1}}\right)^{2}
+\]
+
+\subsection{Code Generation}
 

### 2019-12-30 paper/paper.tex
unicode support
typo

@@
 \usepackage{todonotes}
+\usepackage{textalpha}
 
@@
 We present Soss, a probabilistic programming language embedded in the Julia language.
-Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without reqquiring casual users to worry about such details.
+Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without requiring casual users to worry about such details.
 

### 2019-12-30 paper/paper.tex
BNF: production refinements for 'statements' and 'models'

@@
 
-Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspried by \emph{Stan} \cite{stan:2017}.
+Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspied by \emph{Stan} \cite{stan:2017}.
 
@@
 \begin{bnf*}
-  \bnfprod{model}{\bnfts{@model} \bnfsp \bnfpn{args} \bnfsp \bnfts{begin} \bnfsp \bnfpn{statements} \bnfsp \bnfpn{retn} \bnfsp \bnfts{end}}\\
+  \bnfprod{model}{\bnfts{@model} \bnfsp \bnfpn{args} \bnfsp \bnfts{begin} \bnfsp \bnfpn{statements} \bnfsp \bnfts{end}}\\
   \bnfprod{args}{\bnfes \bnfor \bnfpn{Symbol} \bnfor \bnfpn{Symbol} \bnfts{,} \bnfpn{args}} \\
-  \bnfprod{statements}{\bnfpn{statement} \bnfor \bnfpn{statement} \bnfts{\textbackslash n} \bnfpn{statements}} \\ 
+  \bnfprod{statements}{\bnfpn{statement} \bnfor \bnfpn{statement} \bnfsp \bnftd{Julia Line Sep} \bnfsp \bnfpn{statements} \bnfor \bnfpn{statements} \bnfsp \bnfpn{retn}} \\ 
   \bnfprod{statement}{\bnfpn{assign} \bnfor \bnfpn{sample}} \\

### 2019-12-30 paper/paper.tex
typo
works of GG.jl by Taine Zhao

@@
 
+\todo[inline]{Restructure my writing?}
+
+Besides the purely static code generation via regular macros happening in parsing time, Soss heavily uses a mechanism of "zero-cost" runtime code generation originated by Julia's generated functions \cite{bezanson2012julia}, a.k.a staged functions in the original paper and earlier versions of Julia Programming Language.
+
+The generated functions provide the capability of performing code generation during type inference, generating programs computed by the body of function(a.k.a, generator), once and only once for each combination of argument types.
+
+Further, Julia enables type inference and compiler optimizations equivalent to non-runtime ones in runtime for the callsites of generated functions, hence we gain runtime code generation without losing performance.
+
+To take advantage of this, Soss designs a system to encode sufficient information for generating actual codes, into the types, or objects that can be dispatched like types, which allows the use of generated functions here.
+
+Unfortunately, in current stage, there're some implementation restrictions to Julia's generated functions, and generated functions cannot generate arbitrary Julia codes,
+where some advanced programming constructions are missing, such as generators and function-related stuffs like closures(including functions with no free variables), multiple dispatch, etc.
+
+To address this, we introduced the works of easing the restrictions of Julia generated functions,
+\todo[inline]{cite GG here?}
+which allow us to generate closures from generated functions, and make the programs sufficiently powerful.
+
+Programs that can be generated by Soss are turing complete, as all constructs of Lambda Calculus can be generated.
+
+\begin{lstlisting}[language = Julia]
+  abstract type LCTerm end
+  struct Lam{Arg, Body <: LCTerm} <: LCTerm end
+  struct App{Fn<:LCTerm, Arg<:LCTerm} <: LCTerm end
+  struct Var{N} <: LCTerm end
+
+  cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
+    :($Arg -> $(cg(Term)))
+  cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
+    :($(cg(Fn))($(cg(Arg))))
+  cg(::Type{Var{N}}) where N = N
+
+  function tolc(expr)
+    @match expr begin
+      :($f($arg)) => App{tolc(f), tolc(arg)}
+      :($x => $y) => Lam{x, tolc(y)}
+      a::Symbol => Var{a}
+    end
+  end
+
+  @gg function lc(::Type{Term}) where Term <: LCTerm
+      cg(Term)
+  end
+
+  macro lc(term)
+    lc(tolc(term))
+  end
+
+  x = Core.Box()
+  y = Core.Box()
+  z = Core.Box()
+  f(x) = (x, z)
+  @assert (@lc x => x)(x) === (@lc y => y)(x)
+  @assert (@lc f => x => f(x))(f)(y) === (y, z)
+  \end{lstlisting}
+
 \section{Extending Soss}
@@
 
-Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspied by \emph{Stan} \cite{stan:2017}.
+Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspired by \emph{Stan} \cite{stan:2017}.
 

### 2019-12-30 paper/paper.tex
fix compile error
two-way anova model

@@
 \usepackage{todonotes}
-\usepackage{textalpha}
+% \usepackage{textalpha}
+\usepackage{tikz}
+\usetikzlibrary{bayesnet}
 
@@
 
+\begin{tikzpicture}
+
+    \node[obs] (y) {$y$};
+    \node[det, above=of y] (yhat) {$\hat{y}$};
+    \node[latent, left=of yhat] (epsi) {$\varepsilon_I$};
+    \node[latent, above=of yhat] (alpha) {$\alpha$};
+    \node[latent, right=of yhat] (epsj) {$\varepsilon_J$};
+    \node[latent, left=of alpha] (sigi) {$\sigma_I$};
+    \node[latent, right=of alpha] (sigj) {$\sigma_J$};
+    \node[latent, below=of y] (sigma) {$\sigma$};
+
+    \edge {sigi} {epsi};
+    \edge {sigj} {epsj};
+    \edge {sigma, yhat} {y};
+    \edge {alpha} {sigi};
+    \edge {alpha} {sigj};
+    \edge {epsi,epsj} {yhat};
+
+    \plate {I} {(y)(yhat)(epsi)} {$I$};
+    \plate {J} {(y)(yhat)(epsj)(I.north east)} {$J$};
+
+  \end{tikzpicture}
+
 \begin{lstlisting}[language = Julia]
-m = @model n begin
-    x ~ Normal(0,10) |> iid(n)
-    α ~ Normal()
-    β ~ Normal()
+
+m = @model I,J begin
+    α ~ HalfCauchy()
+    σi ~ HalfNormal(α)
+    σj ~ HalfNormal(α)
+
+    εi ~ Normal(0, σi) |> iid(I)
+    εj ~ Normal(0, σj) |> iid(J)
+    yhat = εz .+ εy'
     σ ~ HalfNormal()
-    yhat = α .+ β .* x
-    y ~ For(n) do j 
-            Normal(yhat[j],σ)
+    y ~ For(I,J) do i,j 
+            Normal(yhat[i,j], σ)
         end
@@
 \[
-  \ell =
-  - 0.5 α^{2} 
-  - 0.5 β^{2}
-  - σ^{2} 
-  + \sum_{\left(j_{1}\right)=1}^{n} \left(- \log{\left(\sigma \right)} 
+    \ell = \beta
+    - 0.5 \alpha^{2} 
+    - 0.5 \beta^{2}
+    - \sigma^{2} 
+    + \sum_{\left(j_{1}\right)=1}^{n} \left(
+        - \log{\left(\sigma \right)} 
         - \frac{\log{\left(\pi \right)}}{2} 
@@
         - \frac{\left({y}_{\left(j_{1}\right)} 
-  - {\hat{y}}_{\left(j_{1}\right)}\right)^{2}}{2 \sigma^{2}}\right) 
-  + \sum_{j_{1}=1}^{n} \left(- 0.005 {x}_{j_{1}}^{2} 
-  - 2.99573227355399 
+        - {\hat{y}}_{\left(j_{1}\right)}\right)^{2}}{2 \sigma^{2}}
+        \right) 
+    + \sum_{j_{1}=1}^{n} \left(
+        - 0.005 {x}_{j_{1}}^{2} 
+        - 3.0 
         - \frac{\log{\left(\pi \right)}}{2} 
-  + \frac{\log{\left(2 \right)}}{2}\right) 
+        + \frac{\log{\left(2 \right)}}{2}
+        \right) 
     - 2 \log{\left(\pi \right)} 

### 2019-12-30 paper/paper.tex
edits

@@
 \usepackage{backnaur}
-\usepackage{jlcode}
+% \usepackage{jlcode}  
 \usepackage{todonotes}
@@
 \usetikzlibrary{bayesnet}
+
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage[utf8x]{inputenc}
+\usepackage{textgreek}
+\usepackage{minted}
 
@@
 
+\begin{figure}[H]
+    \centering
+
+    \begin{subfigure}[b]{0.4\textwidth}
         \begin{tikzpicture}
@@
         \end{tikzpicture}
+        \caption{}
+    \end{subfigure}
+    \hspace{0.1in}
+    \begin{subfigure}[b]{0.4\textwidth}
 
-\begin{lstlisting}[language = Julia]
-
+    \begin{minted}{julia}
     m = @model I,J begin
@@
         end
-\end{lstlisting}
+    \end{minted} 
+    \caption{}
+    \end{subfigure}
+
+    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model}
+    \label{}
+\end{figure}
 
@@
 
-\begin{lstlisting}[language=julia]
+\begin{minted}{julia}
 struct Model{A,B,M} 
@@
 end
-\end{lstlisting}
+\end{minted}
 
@@
 
-\begin{lstlisting}[language = Julia]
+\begin{minted}{julia}
   abstract type LCTerm end
@@
   @assert (@lc f => x => f(x))(f)(y) === (y, z)
-  \end{lstlisting}
+  \end{minted}
 

### 2019-12-30 paper/paper.tex
edits

@@
 \usepackage{todonotes}
+\usepackage{alphabeta}
 % \usepackage{textalpha}
 \usepackage{tikz}
+\usepackage{bookmark}
 \usetikzlibrary{bayesnet}
@@
 \usepackage{subcaption}
-\usepackage[utf8x]{inputenc}
+\usepackage[utf8]{inputenc}
+% \usepackage{unicode-math}
 \usepackage{textgreek}
@@
             \node[det, above=of y] (yhat) {$\hat{y}$};
-            \node[latent, left=of yhat] (epsi) {$\varepsilon_I$};
-            \node[latent, above=of yhat] (alpha) {$\alpha$};
-            \node[latent, right=of yhat] (epsj) {$\varepsilon_J$};
-            \node[latent, left=of alpha] (sigi) {$\sigma_I$};
-            \node[latent, right=of alpha] (sigj) {$\sigma_J$};
+            \node[latent, left=of yhat] (alpha) {$\alpha$};
+            \node[latent, above=of yhat] (lambda) {$\lambda$};
+            \node[latent, right=of yhat] (beta) {$\beta$};
+            \node[latent, left=of lambda] (sigi) {$\sigma_I$};
+            \node[latent, right=of lambda] (sigj) {$\sigma_J$};
             \node[latent, below=of y] (sigma) {$\sigma$};
 
-            \edge {sigi} {epsi};
-            \edge {sigj} {epsj};
+            \edge {sigi} {alpha};
+            \edge {sigj} {beta};
             \edge {sigma, yhat} {y};
-            \edge {alpha} {sigi};
-            \edge {alpha} {sigj};
-            \edge {epsi,epsj} {yhat};
+            \edge {lambda} {sigi};
+            \edge {lambda} {sigj};
+            \edge {alpha,beta} {yhat};
 
-            \plate {I} {(y)(yhat)(epsi)} {$I$};
-            \plate {J} {(y)(yhat)(epsj)(I.north east)} {$J$};
+            \plate {I} {(y)(yhat)(alpha)} {$I$};
+            \plate {J} {(y)(yhat)(beta)(I.north east)} {$J$};
 
@@
 m = @model I,J begin
-        α ~ HalfCauchy()
-        σi ~ HalfNormal(α)
-        σj ~ HalfNormal(α)
+    λ ~ HalfCauchy()
+    σI ~ HalfNormal(λ)
+    σJ ~ HalfNormal(λ)
 
-        εi ~ Normal(0, σi) |> iid(I)
-        εj ~ Normal(0, σj) |> iid(J)
-        yhat = εz .+ εy'
+    α ~ Normal(0, σI) |> iid(I)
+    β ~ Normal(0, σJ) |> iid(J)
+    yhat = α .+ β'
+
     σ ~ HalfNormal()
@@
 
-    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model}
+    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model.}
     \label{}
@@
 \[
-    \ell = \beta
-    - 0.5 \alpha^{2} 
-    - 0.5 \beta^{2}
-    - \sigma^{2} 
-    + \sum_{\left(j_{1}\right)=1}^{n} \left(
-        - \log{\left(\sigma \right)} 
-        - \frac{\log{\left(\pi \right)}}{2} 
-        - \frac{\log{\left(2 \right)}}{2} 
-        - \frac{\left({y}_{\left(j_{1}\right)} 
-        - {\hat{y}}_{\left(j_{1}\right)}\right)^{2}}{2 \sigma^{2}}
-        \right) 
-    + \sum_{j_{1}=1}^{n} \left(
-        - 0.005 {x}_{j_{1}}^{2} 
-        - 3.0 
-        - \frac{\log{\left(\pi \right)}}{2} 
-        + \frac{\log{\left(2 \right)}}{2}
-        \right) 
-    - 2 \log{\left(\pi \right)} 
-    - 1.38629436111989
+\begin{aligned}
+    \ell = 
+       &- 3 \log{\left(2 \right)} 
+        - 5 \log{\left(\pi \right)} 
+    \\ &- σ^{2} 
+        - 4 \log{\left(λ \right)} 
+        - 2 \log{\left(λ^{2} + 1 \right)} 
+        - \frac{σ_I^{2}}{λ^{2}} 
+        - \frac{σ_J^{2}}{λ^{2}}
+    \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
+    \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
+    \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}}\right)
+\end{aligned}
 \]
@@
 \[
+    \begin{aligned}
     \ell =
-  - 3.7 
-  - 4.1 n 
-  - \frac{ \alpha^{2} }{2}
-  - \frac{\beta^2}{2}
-  - \sigma^{2} 
-  - n \log{\left(\sigma \right)} 
-  - 0.005 \sum_{j_{1}=1}^{n} {x}_{j_{1}}^{2} 
-  - \frac{1}{\sigma^2}\sum_{j_{1}=1}^{n} \left({y}_{j_{1}} 
-  - {\hat{y}}_{j_{1}}\right)^{2}
+    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J 
+    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
+        - σ^{2} 
+        - I J \log{σ } 
+    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
+    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
+        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
+        - \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
+    \end{aligned}
 \]

### 2019-12-31 paper/paper.tex
edits

@@
 
+\newmintinline[jl]{julia}{}
 \begin{document}
@@
 \begin{abstract}
-We present Soss, a probabilistic programming language embedded in the Julia language.
+We present Soss, a declarative probabilistic programming language embedded in the Julia language.
 Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without requiring casual users to worry about such details.
@@
 
-Here's some example code. Lots of dead space on the right, maybe want to try to keep code in a multi-column figure
+\section{Model syntax}
 
@@
 
-    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model.}
-    \label{}
+    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model.
+Note the specification that $y$ will later be observed is not part of the model, but is instead given at inference time.}
+    \label{fig:model}
 \end{figure}
 
-\section{Model syntax}
+Figure~\ref{fig:model} shows a plate diagram and Soss implementation of a simplified two-way ANOVA model.
+The model $m$ has \emph{arguments} $I$ and $J$.
+Each line of a model is either a \emph{sample statement} (\jl#v ~ rhs#) or an \emph{assign statement} (\jl#v = rhs#).
+In either case, \jl#v# must be a valid Julia variable name, and \jl#rhs# must be a valid Julia expression.
+For a sample statement, \jl#rhs# should also be ``distribution-like'', supporting \jl#logpdf# and/or \jl#rand# methods (depending on the inference algorithm to be used).
+
+\section{Inference Primitives}
+
+A given inference algorithm can be expressed in terms of a collection of functions on a model.
+For example, we may need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples.
+In Soss, these \emph{inference primitives} are
+
+\begin{figure}
+
+    \begin{minted}{julia}
+        julia> truth = rand(m(I=2,J=3)); pairs(truth)
+        pairs(::NamedTuple) with 10 entries:
+          :I    => 2
+          :J    => 3
+          :yhat => [0.0996722 0.720247 -0.706772; 0.100505 0.72108 -0.705939]
+          :σ    => 0.297288
+          :λ    => 0.523386
+          :σJ   => 0.312794
+          :σI   => 0.0054669
+          :β    => [0.0973136, 0.717889, -0.70913]
+          :α    => [0.00235854, 0.00319107]
+          :y    => [0.386041 0.564963 -0.72177; 0.236898 0.842491 -0.912154]
+    \end{minted}
+
+\end{figure}
 
 \section{Model Transformations}
-
-\subsection{Markov Blankets}
 
@@
 
+\subsection{Markov Blankets}
+
+\begin{minted}{julia}
+julia> markovBlanket(m, :α)
+    @model (I, σI, β, σ, J) begin
+            α ~ Normal(0, σI) |> iid(I)
+            yhat = α .+ β'
+            y ~ For(I, J) do i, j
+                    Normal(yhat[i, j], σ)
+                end
+        end
+\end{minted}
+
 \section{Symbolic Manipulation}
@@
 \subsection{Symbolic Log-density}
-
-\[
-\begin{aligned}
-    \ell = 
-       &- 3 \log{\left(2 \right)} 
-        - 5 \log{\left(\pi \right)} 
-    \\ &- σ^{2} 
-        - 4 \log{\left(λ \right)} 
-        - 2 \log{\left(λ^{2} + 1 \right)} 
-        - \frac{σ_I^{2}}{λ^{2}} 
-        - \frac{σ_J^{2}}{λ^{2}}
-    \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
-    \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
-    \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}}\right)
-\end{aligned}
-\]
-
-\[
-    \begin{aligned}
-    \ell =
-    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J 
-    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
-        - σ^{2} 
-        - I J \log{σ } 
-    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
-    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
-        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
-        - \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
-    \end{aligned}
-\]
 
@@
 
-Authors may provide appendices to accompany their extended abstract submissions. Please note however that reviewers will not be required to comment on material in these appendices.
+\section{Symbolic Simplification}
+Before simplification:
+\[
+\begin{aligned}
+    \ell = 
+       &- 3 \log{\left(2 \right)} 
+        - 5 \log{\left(\pi \right)} 
+    \\ &- σ^{2} 
+        - 4 \log{\left(λ \right)} 
+        - 2 \log{\left(λ^{2} + 1 \right)} 
+        - \frac{σ_I^{2}}{λ^{2}} 
+        - \frac{σ_J^{2}}{λ^{2}}
+    \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
+    \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
+    \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}}\right)
+\end{aligned}
+\]
+
+After simplification:
+\[
+    \begin{aligned}
+    \ell =
+    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J 
+    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
+        - σ^{2} 
+        - I J \log{σ } 
+    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
+    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
+        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
+        - \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
+    \end{aligned}
+\]
+
 \end{document}

### 2019-12-31 paper/paper.tex
edits

@@
 \usepackage{minted} 
+\setminted{fontsize=\footnotesize}
 
@@
 For example, we may need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples.
-In Soss, these \emph{inference primitives} are
+In Soss, these \emph{inference primitives} generate model-specific code at execution time.
 
@@
 \section{Symbolic Simplification}
-Before simplification:
+
+\begin{figure}[t!]
+    \centering
+
+    \begin{subfigure}[t]{0.4\textwidth}
         \[
@@
             - 5 \log{\left(\pi \right)} 
-    \\ &- σ^{2} 
-        - 4 \log{\left(λ \right)} 
+            - σ^{2} 
+        \\ &- 4 \log{\left(λ \right)} 
             - 2 \log{\left(λ^{2} + 1 \right)} 
@@
         \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
-    \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}}\right)
+        \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac    {\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}    }\right)
     \end{aligned}
     \]
-
-After simplification:
+    \caption{Before}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.4\textwidth}
         \[
@@
     \ell =
-    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J 
+    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J - σ^{2} 
     \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
-        - σ^{2} 
         - I J \log{σ } 
@@
         - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
-        - \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
+    \\ &- \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
     \end{aligned}
 \]
+    \captionsetup{skip=27pt}
+    \caption{After}
+    \end{subfigure}
+    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps. Reduction }
+    \label{}
+\end{figure}
+
+Before simplification:
+
+After simplification:
+
+\section{Extended example }
+\todo[inline]{Break this up! This is just a staging area}
+
+\begin{minted}{julia}
+julia> using Revise, Soss, Random
+
+julia> Random.seed!(1);
+
+julia> m = @model I,J begin
+           λ ~ HalfCauchy()
+           σI ~ HalfNormal(λ)
+           σJ ~ HalfNormal(λ)
+           α ~ Normal(0, σI) |> iid(I)
+           β ~ Normal(0, σJ) |> iid(J)
+           yhat = α .+ β'
+           σ ~ HalfNormal()
+           y ~ For(I,J) do i,j 
+                   Normal(yhat[i,j], σ)
+               end
+       end;
+
+julia> truth = rand(m(I=2,J=3)); pairs(truth);
+
+julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));
+
+julia> postpred = predict(m(I=2,J=3), post) |> particles;
+
+julia> postpred.yhat
+2×3 Array{Particles{Float64,1000},2}:
+ 0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
+ 0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37
+
+julia> postpred.y
+2×3 Array{Particles{Float64,1000},2}:
+ 0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
+ 0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61
+
+julia> m.args
+2-element Array{Symbol,1}:
+ :I
+ :J
+
+julia> pairs(m.vals)
+pairs(::NamedTuple) with 1 entry:
+  :yhat => :(α .+ β')
+
+julia> pairs(m.dists)
+pairs(::NamedTuple) with 7 entries:
+  :λ  => :(HalfCauchy())
+  :σI => :(HalfNormal(λ))
+  :σJ => :(HalfNormal(λ))
+  :α  => :(Normal(0, σI) |> iid(I))
+  :β  => :(Normal(0, σJ) |> iid(J))
+  :σ  => :(HalfNormal())
+  :y  => :(For(I, J) do i, j…
+
+\end{minted}
 

### 2019-12-31 paper/paper.tex
example model
edits

@@
 
-There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most in a number of ways...
+There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most alternative in a number of ways...
 
-\section{Model syntax}
+\begin{itemize}
+    \item Model specification in Soss is \emph{declarative}; statements are ordered not by the order entered by the user, but by the partial order given by their variable dependencies.
+    \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distributions within a model.
+    \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, maximizing the flexibility of available inference methods.
+    \item Code generation \texttt{GeneralizedGenerated.jl}
+    \item Model transformations
+    \item Function-like models
+    \item Symbolic simplification
+    \item Extensibility
+\end{itemize}
 
-\begin{figure}[H]
+\section{Modeling in Soss}
+
+Figure~\ref{fig:model} shows a plate diagram and Soss implementation of a simplified two-way ANOVA model.
+The Soss model $m$ has \emph{arguments} $I$ and $J$.
+Each line of a model is either a \emph{sample statement} (\jl#v ~ rhs#) or an \emph{assign statement} (\jl#v = rhs#).
+In either case, \jl#v# must be a valid Julia variable name, and \jl#rhs# must be a valid Julia expression.
+For a sample statement, \jl#rhs# should also be ``distribution-like'', supporting \jl#logpdf# and/or \jl#rand# methods (depending on the inference algorithm to be used).
+
+\begin{figure}[t]
     \centering
 
-    \begin{subfigure}[b]{0.4\textwidth}
-        \begin{tikzpicture}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \[
+    \begin{aligned}
+        \\ i &\in \{1,\cdots,I\}
+        \\ j &\in \{1,\cdots,J\}
+        \\ λ &\sim \text{Cauchy}_+(0,1)
+        \\ \sigma_I &\sim \text{Normal}_+(0,\lambda)
+        \\ \sigma_J &\sim \text{Normal}_+(0,\lambda)
+        \\ \alpha_i &\sim \text{Normal}(0,\sigma_I)
+        \\ \beta_j &\sim \text{Normal}(0,\sigma_J)
+        \\ \hat{y}_{ij} &= \alpha_i + \beta_j
+        \\ y_{ij} &\sim \text{Normal}(\hat{y}_{ij}, \sigma)
+    \end{aligned}
+    \]
+
+    \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \scalebox{0.85}{\begin{tikzpicture}
 
@@
 
-        \end{tikzpicture}
+        \end{tikzpicture}}
         \caption{}
@@
     \hspace{0.1in}
-    \begin{subfigure}[b]{0.4\textwidth}
+    \begin{subfigure}[b]{0.3\textwidth}
 
@@
 
-    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model.
+    \caption{A generative model (a), plate diagram (b), and Soss model (c) for a two-way analysis of variance model.
 Note the specification that $y$ will later be observed is not part of the model, but is instead given at inference time.}
@@
 
-Figure~\ref{fig:model} shows a plate diagram and Soss implementation of a simplified two-way ANOVA model.
-The model $m$ has \emph{arguments} $I$ and $J$.
-Each line of a model is either a \emph{sample statement} (\jl#v ~ rhs#) or an \emph{assign statement} (\jl#v = rhs#).
-In either case, \jl#v# must be a valid Julia variable name, and \jl#rhs# must be a valid Julia expression.
-For a sample statement, \jl#rhs# should also be ``distribution-like'', supporting \jl#logpdf# and/or \jl#rand# methods (depending on the inference algorithm to be used).
-
-\section{Inference Primitives}
+\subsection{Inference Primitives}
 
@@
 
-\begin{figure}
+\subsection{Model Transformations}
 
-    \begin{minted}{julia}
-        julia> truth = rand(m(I=2,J=3)); pairs(truth)
-        pairs(::NamedTuple) with 10 entries:
-          :I    => 2
-          :J    => 3
-          :yhat => [0.0996722 0.720247 -0.706772; 0.100505 0.72108 -0.705939]
-          :σ    => 0.297288
-          :λ    => 0.523386
-          :σJ   => 0.312794
-          :σI   => 0.0054669
-          :β    => [0.0973136, 0.717889, -0.70913]
-          :α    => [0.00235854, 0.00319107]
-          :y    => [0.386041 0.564963 -0.72177; 0.236898 0.842491 -0.912154]
-    \end{minted}
+\subsubsection*{Conditional predictive models}
 
-\end{figure}
+\subsubsection*{Causal interventions (``Do'' operator)}
 
-\section{Model Transformations}
-
-\subsection{Conditional Predictive models}
-
-\subsection{Causal Interventions (``Do'' operator)}
-
-\subsection{Markov Blankets}
+\subsubsection*{Markov blankets}
 
@@
 
-\section{Symbolic Manipulation}
+\subsection{Symbolic Manipulation}
 
-\subsection{Symbolic Log-density}
+\subsubsection*{Symbolic Log-density}
 
-\subsection{Code Generation}
+\subsubsection*{Code Generation}
 

### 2019-12-31 paper/paper.tex
edits

@@
   %\streetaddress{43 Vassar St}
-  %\city{Cambridge}
-  %\state{MA}
+  \city{Seattle}
+  \state{WA}
   %\postcode{02139}
@@
     \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distributions within a model.
+    \item Soss models are \emph{function-like}; abstractly, a model may be considered a function from its arguments to a joint distribution.
+In particular, specification of ``observed variables'' is separated from model definition.
     \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, maximizing the flexibility of available inference methods.
-    \item Code generation \texttt{GeneralizedGenerated.jl}
+    \item Soss uses \emph{staged compilation} for inference primitives like \jl#rand# and \jl#logpdf#, via novel \emph{generalized generated functions} from \texttt{GeneralizedGenerated.jl}.
     \item Model transformations
-    \item Function-like models
     \item Symbolic simplification
@@
 \end{itemize}
+
+Finally, Soss addresses the age-old challenge of naming things in computer science via Cockney rhyming slang (``sauce pan'' rhymes with ``Stan''). {\bf S}oss is {\bf o}pen-{\bf s}ource {\bf s}oftware.
 

### 2019-12-31 paper/paper.tex
edits

@@
     \item Model specification in Soss is \emph{declarative}; statements are ordered not by the order entered by the user, but by the partial order given by their variable dependencies.
+    \item Soss models are typically expressed in terms of \emph{inference combinators} (like ``\jl#For#'') that combine distributions in some way to arrive at a new distribution.
     \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distributions within a model.
@@
 In particular, specification of ``observed variables'' is separated from model definition.
-    \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, maximizing the flexibility of available inference methods.
+    \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, keeping the internal representation close to that given by the user for maximum flexibility.
     \item Soss uses \emph{staged compilation} for inference primitives like \jl#rand# and \jl#logpdf#, via novel \emph{generalized generated functions} from \texttt{GeneralizedGenerated.jl}.
-    \item Model transformations
-    \item Symbolic simplification
-    \item Extensibility
+    \item Soss supports \emph{model transformations}, functions that take a model and return another model.
+    \item Soss supports \emph{symbolic simplifcation}, making it easy to inspect or manipulate a symbolic representation of the log-density, or to use it to generate optimized code.
+    \item Soss is \emph{extensible}; users can define new inference primitives or model transformations externally, and use them as if they had been included in Soss.
 \end{itemize}
 
-Finally, Soss addresses the age-old challenge of naming things in computer science via Cockney rhyming slang (``sauce pan'' rhymes with ``Stan''). {\bf S}oss is {\bf o}pen-{\bf s}ource {\bf s}oftware.
+Finally, Soss addresses the well-known challenge of naming things in computer science via Cockney rhyming slang (``sauce pan'' rhymes with ``Stan'').
+{\bf S}oss is {\bf o}pen-{\bf s}ource {\bf s}oftware.
 

### 2019-12-31 paper/paper.tex
edits

@@
 
-There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most alternative in a number of ways...
+There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most alternatives in a number of ways:
 
 \begin{itemize}
-    \item Model specification in Soss is \emph{declarative}; statements are ordered not by the order entered by the user, but by the partial order given by their variable dependencies.
+    \item Model specification in Soss is \emph{declarative}; statements are represented not by the order entered by the user, but by the partial order given by their variable dependencies.
     \item Soss models are typically expressed in terms of \emph{inference combinators} (like ``\jl#For#'') that combine distributions in some way to arrive at a new distribution.
@@
 
+\subsection{Joint Distributions}
+
 \subsection{Inference Primitives}
@@
 In Soss, these \emph{inference primitives} generate model-specific code at execution time.
+
+At this writing, the following inference primitive are available:
+
+\begin{description}  
+    \item[\texttt{rand}] draws a random sample from a given joint distribution 
+    \item[\texttt{particles}] uses \emph{systematic sampling}
+    \item[\texttt{logpdf}]
+    \item[\texttt{symlogpdf}] 
+    \item[\texttt{weightedSample}] 
+    \item[\texttt{xform}] 
+\end{description}
 

### 2020-01-01 paper/paper.tex
edits

@@
 
+\todo[inline]{citation for MonteCarloMeasurements.jl}
+
 \begin{description}  
-    \item[\texttt{rand}] draws a random sample from a given joint distribution 
-    \item[\texttt{particles}] uses \emph{systematic sampling}
-    \item[\texttt{logpdf}]
+    \item[\texttt{rand}] draws a random sample from a given joint distribution.
+    \item[\texttt{particles}] draws a \emph{systematic sample} from a given joint distribution, using \texttt{MonteCarloMeasurements.jl}.
+    \item[\texttt{logpdf}] takes a joint distribution and a point in its sample space, and returns the log-density.
     \item[\texttt{symlogpdf}] 

### 2020-01-01 paper/paper.tex
pdf
edits

@@
 
-Finally, Soss addresses the well-known challenge of naming things in computer science via Cockney rhyming slang (``sauce pan'' rhymes with ``Stan'').
+Finally, Soss uses Cockney rhyming slang to address the well-known challenge of naming things in computer science (``sauce pan'' rhymes with ``Stan'').
 {\bf S}oss is {\bf o}pen-{\bf s}ource {\bf s}oftware.
@@
 \subsection{Joint Distributions}
+
+A Soss model is \emph{function-like}, taking values for its arguments and returning a \emph{joint distribution}.
 

### 2020-01-01 paper/paper.tex
thautwarm edits

--- 2019-12-31 paper/paper.tex
@@
   %\streetaddress{43 Vassar St}
-  \city{Seattle}
-  \state{WA}
+  %\city{Cambridge}
+  %\state{MA}
   %\postcode{02139}
@@
 
-There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most alternatives in a number of ways:
+There are many approaches to building a Probabilistic Programming Language (``PPL''). Soss is distinct from most in a number of ways...
 
-\begin{itemize}
-    \item Model specification in Soss is \emph{declarative}; statements are represented not by the order entered by the user, but by the partial order given by their variable dependencies.
-    \item Soss models are typically expressed in terms of \emph{inference combinators} (like ``\jl#For#'') that combine distributions in some way to arrive at a new distribution.
-    \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distributions within a model.
-    \item Soss models are \emph{function-like}; abstractly, a model may be considered a function from its arguments to a joint distribution.
-In particular, specification of ``observed variables'' is separated from model definition.
-    \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, keeping the internal representation close to that given by the user for maximum flexibility.
-    \item Soss uses \emph{staged compilation} for inference primitives like \jl#rand# and \jl#logpdf#, via novel \emph{generalized generated functions} from \texttt{GeneralizedGenerated.jl}.
-    \item Soss supports \emph{model transformations}, functions that take a model and return another model.
-    \item Soss supports \emph{symbolic simplifcation}, making it easy to inspect or manipulate a symbolic representation of the log-density, or to use it to generate optimized code.
-    \item Soss is \emph{extensible}; users can define new inference primitives or model transformations externally, and use them as if they had been included in Soss.
-\end{itemize}
+\section{Model syntax}
 
-Finally, Soss addresses the well-known challenge of naming things in computer science via Cockney rhyming slang (``sauce pan'' rhymes with ``Stan'').
-{\bf S}oss is {\bf o}pen-{\bf s}ource {\bf s}oftware.
-
-\section{Modeling in Soss}
-
-Figure~\ref{fig:model} shows a plate diagram and Soss implementation of a simplified two-way ANOVA model.
-The Soss model $m$ has \emph{arguments} $I$ and $J$.
-Each line of a model is either a \emph{sample statement} (\jl#v ~ rhs#) or an \emph{assign statement} (\jl#v = rhs#).
-In either case, \jl#v# must be a valid Julia variable name, and \jl#rhs# must be a valid Julia expression.
-For a sample statement, \jl#rhs# should also be ``distribution-like'', supporting \jl#logpdf# and/or \jl#rand# methods (depending on the inference algorithm to be used).
-
-\begin{figure}[t]
+\begin{figure}[H]
     \centering
 
-    \begin{subfigure}[b]{0.3\textwidth}
-        \[
-    \begin{aligned}
-        \\ i &\in \{1,\cdots,I\}
-        \\ j &\in \{1,\cdots,J\}
-        \\ λ &\sim \text{Cauchy}_+(0,1)
-        \\ \sigma_I &\sim \text{Normal}_+(0,\lambda)
-        \\ \sigma_J &\sim \text{Normal}_+(0,\lambda)
-        \\ \alpha_i &\sim \text{Normal}(0,\sigma_I)
-        \\ \beta_j &\sim \text{Normal}(0,\sigma_J)
-        \\ \hat{y}_{ij} &= \alpha_i + \beta_j
-        \\ y_{ij} &\sim \text{Normal}(\hat{y}_{ij}, \sigma)
-    \end{aligned}
-    \]
-
-    \caption{}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \scalebox{0.85}{\begin{tikzpicture}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \begin{tikzpicture}
 
@@
 
-        \end{tikzpicture}}
+        \end{tikzpicture}
         \caption{}
@@
     \hspace{0.1in}
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.4\textwidth}
 
@@
 
-    \caption{A generative model (a), plate diagram (b), and Soss model (c) for a two-way analysis of variance model.
+    \caption{A plate diagram (a) and Soss encoding (b) for a two-way analysis of variance model.
 Note the specification that $y$ will later be observed is not part of the model, but is instead given at inference time.}
@@
 
-\subsection{Joint Distributions}
+Figure~\ref{fig:model} shows a plate diagram and Soss implementation of a simplified two-way ANOVA model.
+The model $m$ has \emph{arguments} $I$ and $J$.
+Each line of a model is either a \emph{sample statement} (\jl#v ~ rhs#) or an \emph{assign statement} (\jl#v = rhs#).
+In either case, \jl#v# must be a valid Julia variable name, and \jl#rhs# must be a valid Julia expression.
+For a sample statement, \jl#rhs# should also be ``distribution-like'', supporting \jl#logpdf# and/or \jl#rand# methods (depending on the inference algorithm to be used).
 
-\subsection{Inference Primitives}
+\section{Inference Primitives}
 
@@
 
-At this writing, the following inference primitive are available:
+\begin{figure}
 
-\begin{description}  
-    \item[\texttt{rand}] draws a random sample from a given joint distribution 
-    \item[\texttt{particles}] uses \emph{systematic sampling}
-    \item[\texttt{logpdf}]
-    \item[\texttt{symlogpdf}] 
-    \item[\texttt{weightedSample}] 
-    \item[\texttt{xform}] 
-\end{description}
+    \begin{minted}{julia}
+        julia> truth = rand(m(I=2,J=3)); pairs(truth)
+        pairs(::NamedTuple) with 10 entries:
+          :I    => 2
+          :J    => 3
+          :yhat => [0.0996722 0.720247 -0.706772; 0.100505 0.72108 -0.705939]
+          :σ    => 0.297288
+          :λ    => 0.523386
+          :σJ   => 0.312794
+          :σI   => 0.0054669
+          :β    => [0.0973136, 0.717889, -0.70913]
+          :α    => [0.00235854, 0.00319107]
+          :y    => [0.386041 0.564963 -0.72177; 0.236898 0.842491 -0.912154]
+    \end{minted}
 
-\subsection{Model Transformations}
+\end{figure}
 
-\subsubsection*{Conditional predictive models}
+\section{Model Transformations}
 
-\subsubsection*{Causal interventions (``Do'' operator)}
+\subsection{Conditional Predictive models}
 
-\subsubsection*{Markov blankets}
+\subsection{Causal Interventions (``Do'' operator)}
+
+\subsection{Markov Blankets}
 
@@
 
-\subsection{Symbolic Manipulation}
+\section{Symbolic Manipulation}
 
-\subsubsection*{Symbolic Log-density}
+\subsection{Symbolic Log-density}
 
-\subsubsection*{Code Generation}
+\subsection{Code Generation}
 
@@
 
-Besides the purely static code generation via regular macros happening in parsing time, Soss heavily uses a mechanism of "zero-cost" runtime code generation originated by Julia's generated functions \cite{bezanson2012julia}, a.k.a staged functions in the original paper and earlier versions of Julia Programming Language.
+Besides the purely static code generation via regular macros happening in parsing time, Soss heavily uses a mechanism of "zero-cost" runtime code generation originated by Julia's generated functions, a.k.a staged functions in the original paper and earlier versions of Julia Programming Language \cite{bezanson2012julia}.
 
@@
 
-To address this, we introduced the works of easing the restrictions of Julia generated functions,
-\todo[inline]{cite GG here?}
-which allow us to generate closures from generated functions, and make the programs sufficiently powerful.
+To address this, we ease the restrictions of Julia generated functions by leveraging GeneralizedGenerated \cite{taine_zhao_2020_3596233} to,
+which allows us to generate closures from generated functions.
 
@@
 
-  cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
-    :($Arg -> $(cg(Term)))
-  cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
-    :($(cg(Fn))($(cg(Arg))))
+  cg(::Type{Lam{Arg, Term}}) where {Arg, Term} =  :($Arg -> $(cg(Term)))
+  cg(::Type{App{Fn, Arg}}) where {Fn, Arg} =  :($(cg(Fn))($(cg(Arg))))
   cg(::Type{Var{N}}) where N = N
 
-  function tolc(expr)
-    @match expr begin
-      :($f($arg)) => App{tolc(f), tolc(arg)}
-      :($x => $y) => Lam{x, tolc(y)}
-      a::Symbol => Var{a}
-    end
-  end
+  @gg lc(::Type{Term}) where Term <: LCTerm = cg(Term)
+\end{minted} 
+  % tolc(expr::Expr) =  @match expr begin
+  %                       :($f($arg)) => App{tolc(f), tolc(arg)}
+  %                       :($x => $y) => Lam{x, tolc(y)}
+  %                     end
+  % tolc(sym::Symbol) => Var{sym}
+  % macro lc(term)
+  %   lc(tolc(term))
+  % end
 
-  @gg function lc(::Type{Term}) where Term <: LCTerm
-      cg(Term)
-  end
-
-  macro lc(term)
-    lc(tolc(term))
-  end
-
-  x = Core.Box()
-  y = Core.Box()
-  z = Core.Box()
-  f(x) = (x, z)
-  @assert (@lc x => x)(x) === (@lc y => y)(x)
-  @assert (@lc f => x => f(x))(f)(y) === (y, z)
-  \end{minted}
+  % x = Core.Box()
+  % y = Core.Box()
+  % z = Core.Box()
+  % f(x) = (x, z)
+  % @assert (@lc x => x)(x) === (@lc y => y)(x)
+  % @assert (@lc f => x => f(x))(f)(y) === (y, z)
 

### 2020-01-01 paper/paper.tex
edits

--- 2020-01-01 paper/paper.tex
@@
     \item Soss models are typically expressed in terms of \emph{inference combinators} (like ``\jl#For#'') that combine distributions in some way to arrive at a new distribution.
-    \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distributions within a model.
+    \item Soss models are \emph{first-class}; models can be passed as arguments to other models, or used in place of distribution functions like \jl#Normal# within a model.
     \item Soss models are \emph{function-like}; abstractly, a model may be considered a function from its arguments to a joint distribution.
@@
     \item Values and distributions in Soss are represented internally as \emph{abstract syntax trees}, keeping the internal representation close to that given by the user for maximum flexibility.
-    \item Soss uses \emph{staged compilation} for inference primitives like \jl#rand# and \jl#logpdf#, via novel \emph{generalized generated functions} from \texttt{GeneralizedGenerated.jl}.
+    \item Soss uses \emph{staged compilation} for inference primitives like \jl#rand# and \jl#logpdf#, via novel \emph{generalized generated functions} from \jl#GeneralizedGenerated.jl#.
     \item Soss supports \emph{model transformations}, functions that take a model and return another model.
@@
 A Soss model is \emph{function-like}, taking values for its arguments and returning a \emph{joint distribution}.
+For example, given \jl#m# from Figure~\ref{fig:model}, we can build the joint distribution \jl#d=m(I=2, J=3)#.
+This supports the usual \jl#rand# and \jl#logpdf# functions, as well as some others provided by Soss described in the following sections.
+For example,\footnote{This and future examples are edited for space, for example rounding floating point values.}
+
+\begin{minted}{julia}
+julia> d = m(I=2, J=3); rand(d) # or rand(m(I=2, J=3))
+(I = 2, J = 3, yhat = [-0.52 -1.5 0.0037; -0.81 -1.85 -0.28], σ = 2.5, λ = 1.3, σJ = 0.72, σI = 0.41, 
+ β = [-0.54, -1.5, -0.023], α = [0.026, -0.26], y = [-0.45 -2.0 2.2; -3.4 -3.9 2.3])
+\end{minted}
 

### 2020-01-01 paper/paper.tex
edits

@@
 This supports the usual \jl#rand# and \jl#logpdf# functions, as well as some others provided by Soss described in the following sections.
-For example,\footnote{This and future examples are edited for space, for example rounding floating point values.}
+For example,\footnote{This and future outputs are lightly edited for space, for example rounding floating point values.}
 
 \begin{minted}{julia}
-julia> d = m(I=2, J=3); rand(d) # or rand(m(I=2, J=3))
+julia> d = m(I=2, J=3); rand(d) # or just rand(m(I=2, J=3))
 (I = 2, J = 3, yhat = [-0.52 -1.5 0.0037; -0.81 -1.85 -0.28], σ = 2.5, λ = 1.3, σJ = 0.72, σI = 0.41, 
@@
 
-A given inference algorithm can be expressed in terms of a collection of functions on a model.
-For example, we may need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples.
+A given inference algorithm can be expressed in terms of a collection of functions on a model or joint distribution.
+For example, we often need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples.
 In Soss, these \emph{inference primitives} generate model-specific code at execution time.
@@
 
-\todo[inline]{citation for MonteCarloMeasurements.jl}
+\begin{description}  
+    \item[\texttt{rand}] draws a random point or sample from a given distribution.
+    \item[\texttt{particles}] draws a \emph{systematic sample}, as described in \cite{Douc2005}.
+    \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
+    \item[\texttt{xform}] takes a distribution and some known data, and returns a bijection from $\mathbb{R}^n$ to the space of unknowns.
+This is useful for algorithms like Hamiltonian Monte Carlo \cite{Neal2011}.
+    \item[\texttt{symlogpdf}] 
+    \item[\texttt{codegen}]  
+\end{description}
 
-\begin{description}  
-    \item[\texttt{rand}] draws a random sample from a given joint distribution.
-    \item[\texttt{particles}] draws a \emph{systematic sample} from a given joint distribution, using \texttt{MonteCarloMeasurements.jl}.
-    \item[\texttt{logpdf}] takes a joint distribution and a point in its sample space, and returns the log-density.
-    \item[\texttt{symlogpdf}] 
-    \item[\texttt{weightedSample}] 
-    \item[\texttt{xform}] 
-\end{description}
+In the above, a \emph{distribution} refers to a joint distribution, while \emph{point} refers to a named tuple in the space space of a given distribution.
+
+\todo[inline]{cite MonteCarloMeasurements.jl and TransformVariables.jl}
+
+Users may add more
 

### 2020-01-02 paper/paper.tex
edits

@@
 A given inference algorithm can be expressed in terms of a collection of functions on a model or joint distribution.
-For example, we often need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples.
+For example, we often need to evaluate the log-density or its gradient  or draw random samples.
 In Soss, these \emph{inference primitives} generate model-specific code at execution time.
@@
 This is useful for algorithms like Hamiltonian Monte Carlo \cite{Neal2011}.
-    \item[\texttt{symlogpdf}] 
-    \item[\texttt{codegen}]  
+    \item[\texttt{symlogpdf}] takes a model, and returns a SymPy \cite{10.7717/peerj-cs.103} representation of the log-density, with some additional Julia-based simplifications.
+    \item[\texttt{codegen}] calls \jl#symlogpdf#, and uses the result to generate a more efficient version of \jl#logpdf#.
 \end{description}
@@
 
+\subsection{Functions on Models}
+
+arguments
+sampled
+assigned
+parameters
+variables
+
+toposort
+poset
+digraph
+
 \subsection{Model Transformations}
 
-\subsubsection*{Conditional predictive models}
-
-\subsubsection*{Causal interventions (``Do'' operator)}
-
-\subsubsection*{Markov blankets}
-
-\begin{minted}{julia}
-julia> markovBlanket(m, :α)
-    @model (I, σI, β, σ, J) begin
-            α ~ Normal(0, σI) |> iid(I)
-            yhat = α .+ β'
-            y ~ For(I, J) do i, j
-                    Normal(yhat[i, j], σ)
-                end
-        end 
-\end{minted}
-
-\subsection{Symbolic Manipulation}
-
-\subsubsection*{Symbolic Log-density}
-
-\subsubsection*{Code Generation}
+\begin{description}
+    \item[\texttt{Do}] implements Pearl's {\footnotesize\texttt{do}} operator for causal intervention \cite{Pearl1995}.\footnote{Our {\footnotesize\texttt{Do}} operator is capitalized because {\footnotesize\texttt{do}} is a keyword in Julia.}
+    \item[\texttt{predictive}] is like \jl#Do#, but only returns the strongly-connected components containing the specified variables.
+This is useful for sampling from the posterior predictive distribution.
+    \item[\texttt{markovBlanket}]  takes a model and a variable, and returns a model representing the Markov blanket at the given variable.
+\end{description}
 
@@
 
-\begin{minted}{julia}
-  abstract type LCTerm end
-  struct Lam{Arg, Body <: LCTerm} <: LCTerm end
-  struct App{Fn<:LCTerm, Arg<:LCTerm} <: LCTerm end
-  struct Var{N} <: LCTerm end
+% \begin{minted}{julia}
+%   abstract type LCTerm end
+%   struct Lam{Arg, Body <: LCTerm} <: LCTerm end
+%   struct App{Fn<:LCTerm, Arg<:LCTerm} <: LCTerm end
+%   struct Var{N} <: LCTerm end
 
-  cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
-    :($Arg -> $(cg(Term)))
-  cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
-    :($(cg(Fn))($(cg(Arg))))
-  cg(::Type{Var{N}}) where N = N
+%   cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
+%     :($Arg -> $(cg(Term)))
+%   cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
+%     :($(cg(Fn))($(cg(Arg))))
+%   cg(::Type{Var{N}}) where N = N
 
-  function tolc(expr)
-    @match expr begin
-      :($f($arg)) => App{tolc(f), tolc(arg)}
-      :($x => $y) => Lam{x, tolc(y)}
-      a::Symbol => Var{a}
-    end
-  end
+%   function tolc(expr)
+%     @match expr begin
+%       :($f($arg)) => App{tolc(f), tolc(arg)}
+%       :($x => $y) => Lam{x, tolc(y)}
+%       a::Symbol => Var{a}
+%     end
+%   end
 
-  @gg function lc(::Type{Term}) where Term <: LCTerm
-      cg(Term)
-  end
+%   @gg function lc(::Type{Term}) where Term <: LCTerm
+%       cg(Term)
+%   end
 
-  macro lc(term)
-    lc(tolc(term))
-  end
+%   macro lc(term)
+%     lc(tolc(term))
+%   end
 
-  x = Core.Box()
-  y = Core.Box()
-  z = Core.Box()
-  f(x) = (x, z)
-  @assert (@lc x => x)(x) === (@lc y => y)(x)
-  @assert (@lc f => x => f(x))(f)(y) === (y, z)
-  \end{minted}
+%   x = Core.Box()
+%   y = Core.Box()
+%   z = Core.Box()
+%   f(x) = (x, z)
+%   @assert (@lc x => x)(x) === (@lc y => y)(x)
+%   @assert (@lc f => x => f(x))(f)(y) === (y, z)
+%   \end{minted}
 

### 2020-01-02 paper/paper.tex
edits

@@
 
-Users may add more
+The set of inference primitives is not fixed, but can be extended by users. New primitives can be added in outside packages, and does not require making changes to the Soss codebase.
 
@@
 
-arguments
-sampled
-assigned
-parameters
-variables
+Soss includes several functions for querying sets of variables associated with a model.
+As described above, models are function-like, with the inputs given by {\bf \texttt{arguments}}.
+In the body of a model every statement declares that some variable is either {\bf \texttt{sampled}} or {\bf \texttt{assigned}}.
+The sampled and assigned variables together comprise the {\bf \texttt{parameters}}, and all of the above are {\bf \texttt{variables}}.
+Each of these is a Soss function that takes a model and returns a vector of distinct Julia symbols.
 
-toposort
-poset
-digraph
+Soss ignores the ordering of statements given by the user, instead working in terms of the DAG ordering.
+You can access this graph with {\bf \texttt{digraph}}, or use {\bf \texttt{poset}} to get the partial order.
+Finally, {\bf \texttt{toposort}} returns a topologically-ordered vector of the parameters.
+Each of these three functions takes a model as input.
+
+\todo[inline]{Cite SimplePosets and SimpleGraphs}
 

### 2020-01-02 paper/paper.tex
edits

@@
 
-In the above, a \emph{distribution} refers to a joint distribution, while \emph{point} refers to a named tuple in the space space of a given distribution.
+In the above, a \emph{distribution} refers to a joint distribution, while \emph{point} refers to a named tuple in the space space of a distribution.
 
@@
 The set of inference primitives is not fixed, but can be extended by users. New primitives can be added in outside packages, and does not require making changes to the Soss codebase.
+
+\subsection{Inference}
+
+\jl#DynamicHMC.jl# \cite{tamas_k_papp_2019_3590018}
 
@@
 
-\todo[inline]{Cite SimplePosets and SimpleGraphs}
+Soss represents and manipulates variable dependencies using \jl#SimpleGraphs#, \jl#SimplePosets#, and \jl#SimplePartitions# from \cite{SimpleWorld}.
 

### 2020-01-03 paper/paper.tex
edits

@@
 \setminted{fontsize=\footnotesize}
+\usepackage{newunicodechar}
+\newunicodechar{ℓ}{\ensuremath{\ell}}
 
@@
 
-\jl#DynamicHMC.jl# \cite{tamas_k_papp_2019_3590018}
+Together with automatic differentiation, the above inference primitives enable a wide variety of inference algorithms.
+Rather than building such algorithms specific to Soss, we defer this to external packages.
+So far, our focus has been Hamiltonian Monte Carlo \cite{Neal2011}, as implemented in \jl#DynamicHMC.jl# \cite{tamas_k_papp_2019_3590018} and \jl#AdvancedHMC.jl# \cite{ge2018t}; Soss has interfaces to these in the {\bf \texttt{dynamicHMC}} and {\bf \texttt{advancedHMC}} functions, respectively.
 
@@
 \end{minted}
-
-\section{Performance}
-
-\todo[inline]{Let's implement models from \url{https://statisticalrethinkingjulia.github.io/MCMCBenchmarks.jl/latest/benchmarks/} for benchmark comparisons}
-
-\todo[inline]{Can Soss really generate arbitrary code?}
 
@@
     \]
-    \caption{Before}
+    \caption{Direct implementation}
     \end{subfigure}
@@
     \captionsetup{skip=27pt}
-    \caption{After}
+    \caption{After symbolic simplification}
     \end{subfigure}
-    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps. Reduction }
+    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps.}
     \label{}
@@
 
-Before simplification:
-
-After simplification:
+\begin{figure}[t]
+\centering
+\begin{subfigure}[t]{0.48\textwidth}
+\begin{minted}{julia}
+julia> Soss.sourceRand()(m)
+quote
+    σ = rand(HalfNormal())
+    λ = rand(HalfCauchy())
+    σJ = rand(HalfNormal(λ))
+    σI = rand(HalfNormal(λ))
+    β = rand(iid(J, Normal(0, σJ)))
+    α = rand(iid(I, Normal(0, σI)))
+    yhat = α .+ β'
+    y = rand(For(I, J) do i, j
+                Normal(yhat[i, j], σ)
+            end)
+    (I=I, J=J, yhat=yhat, λ=λ, σI=σI, 
+    σJ=σJ, α=α, β=β, σ=σ, y=y) 
+end
+\end{minted}
+\caption{\footnotesize{\texttt{rand}}}
+\end{subfigure}
+\begin{subfigure}[t]{0.48\textwidth}
+\begin{minted}{julia}
+julia> Soss.sourceLogpdf()(m)
+quote
+    _ℓ = 0.0
+    _ℓ += logpdf(HalfNormal(), σ)
+    _ℓ += logpdf(HalfCauchy(), λ)
+    _ℓ += logpdf(HalfNormal(λ), σJ)
+    _ℓ += logpdf(HalfNormal(λ), σI)
+    _ℓ += logpdf(Normal(0, σJ) |> iid(J), β)
+    _ℓ += logpdf(Normal(0, σI) |> iid(I), α)
+    yhat = α .+ β'
+    _ℓ += logpdf(For(I, J) do i, j
+                Normal(yhat[i, j], σ)
+            end, y)
+    return _ℓ
+end
+\end{minted}
+\caption{\footnotesize{\texttt{logpdf}}}
+\end{subfigure}
+\caption{Generated code when calling \footnotesize{\texttt{rand(m)}} and \footnotesize{\texttt{logpdf(m,x)}}.} 
+\label{} 
+\end{figure}
 

### 2020-01-04 paper/paper.tex

@@
 
-%   cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
-%     :($Arg -> $(cg(Term)))
-%   cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
-%     :($(cg(Fn))($(cg(Arg))))
+%   cg(::Type{Lam{Arg, Term}}) where {Arg, Term} =  :($Arg -> $(cg(Term)))
+%   cg(::Type{App{Fn, Arg}}) where {Fn, Arg} =  :($(cg(Fn))($(cg(Arg))))
 %   cg(::Type{Var{N}}) where N = N
 
-%   function tolc(expr)
-%     @match expr begin
+%   @gg lc(::Type{Term}) where Term <: LCTerm = cg(Term)
+% \end{minted} 
+  % tolc(expr::Expr) =  @match expr begin
   %                       :($f($arg)) => App{tolc(f), tolc(arg)}
   %                       :($x => $y) => Lam{x, tolc(y)}
-%       a::Symbol => Var{a}
   %                     end
-%   end
-
-%   @gg function lc(::Type{Term}) where Term <: LCTerm
-%       cg(Term)
-%   end
-
+  % tolc(sym::Symbol) => Var{sym}
   % macro lc(term)
@@
   % @assert (@lc f => x => f(x))(f)(y) === (y, z)
-%   \end{minted}
 

### 2020-01-05 paper/paper.tex
edits

--- 2020-01-03 paper/paper.tex
@@
 
-\newmintinline[jl]{julia}{}
+\newmintinline[jl]{julia1}{}
 \begin{document}
@@
 
-    \begin{minted}{julia}
+    \begin{minted}{julia1}
 m = @model I,J begin
@@
 
-\begin{minted}{julia}
+\begin{minted}{julia1}
 julia> d = m(I=2, J=3); rand(d) # or just rand(m(I=2, J=3))
@@
 
-\begin{minted}{julia}
+Soss's inference primitives generate code on demand using the \jl#GeneralizedGenerated.jl# library, which introduces a \jl#@gg# macro that works as an extension of Julia's built-in \jl#@generated# macro for \emph{generated functions} \cite{bezanson2012julia}.
+Similarly to \jl#@generated#, code generated using a \jl#@gg# function cannot depend on the values of its arguments, but only on their types.
+
+This is problematic, since we clearly need code generation to depend on the model itself. To address this, \jl#GeneralizedGenerated# includes facilities for lifting Julia values to the type level.
+
+Using this, the type of a Soss model is \jl#Model{A,B,M}#, where 
+\begin{itemize}
+    \item \jl#A# gives the type of the \emph{arguments} (\jl#A=NamedTuple{T}# for some tuple of symbols \jl#T#),
+    \item \jl#B# is a type-level representation of the \emph{body}, and
+    \item \jl#M# is a type-level representation of the \emph{module} in which the model was defined.
+\end{itemize}
+
+In this, the ``body'' is the \jl#begin...end# following the (possibly empty) arguments of a model. Knowing the module is important in case values from that module are referenced in the model.
+
+\begin{figure}[t]
+    \begin{subfigure}[b]{0.48\textwidth}
+        \begin{minted}[autogobble]{julia1}
         struct Model{A,B,M} 
@@
         end
+
+        julia> m.dists
+        ( λ = :(HalfCauchy())
+        , σI = :(HalfNormal(λ))
+        , σJ = :(HalfNormal(λ))
+        , α = :(Normal(0, σI) |> iid(I))
+        , β = :(Normal(0, σJ) |> iid(J))
+        , σ = :(HalfNormal())
+        , y = :(For(I, J) do i, j
+                Normal(yhat[i, j], σ)
+                end))
         \end{minted}
 
-The ability to generate arbitrary  Julia code makes it difficult to measure performance in Soss.
+    \end{subfigure}    
+    \begin{subfigure}[b]{0.48\textwidth}
+        \begin{minted}[autogobble]{julia1}
+            julia> m.vals
+            (yhat = :(α .+ β'),)
+
+            julia> typeof(m)
+            Model{NamedTuple{(:I, :J),T} where T<:Tuple,
+                TypeEncoding(begin
+                    σ ~ HalfNormal()
+                    λ ~ HalfCauchy()
+                    σJ ~ HalfNormal(λ)
+                    σI ~ HalfNormal(λ)
+                    β ~ Normal(0, σJ) |> iid(J)
+                    α ~ Normal(0, σI) |> iid(I)
+                    yhat = α .+ β'
+                    y ~ For(I, J) do i, j
+                            Normal(yhat[i, j], σ)
+                        end
+                end),TypeEncoding(Main)}
+        \end{minted}
+    \end{subfigure}    
+    \caption{The definition of the \texttt{Model} type, and some details of the \texttt{m} model from Figure \ref{fig:model}.}
+\end{figure}
+
+The flexibility of code generation in Soss makes it difficult to measure performance.
 When a performance bottleneck is found, developers or users can ask Soss to generate something different.
@@
 
-% \begin{minted}{julia}
+% \begin{minted}{julia1}
 %   abstract type LCTerm end
@@
 \begin{subfigure}[t]{0.48\textwidth}
-\begin{minted}{julia}
+\begin{minted}{julia1}
 julia> Soss.sourceRand()(m)
@@
 \begin{subfigure}[t]{0.48\textwidth}
-\begin{minted}{julia}
+\begin{minted}{julia1}
 julia> Soss.sourceLogpdf()(m)
@@
 
-\begin{minted}{julia}
+\begin{minted}{julia1}
 julia> using Revise, Soss, Random

### 2020-01-05 paper/paper.tex
edits

@@
 
-This is problematic, since we clearly need code generation to depend on the model itself. To address this, \jl#GeneralizedGenerated# includes facilities for lifting Julia values to the type level.
+This \emph{staged compilation} introduces some additional overhead on the first call on a given argument type, due to generation and compilation of specialized code.
+Subsequent calls on this same type have no such overhead, and generally benefit from specialization.
+To acheive a net gain from this requires making enough function calls to offset the cost of code generation.
+This makes the approach a natural fit for the PPL domain, where long-running tight loops are very common.
 
-Using this, the type of a Soss model is \jl#Model{A,B,M}#, where 
+However, the requirement that the generated code only depends on the types presents a challenge, since we clearly need it to depend on the model itself.
+To address this, \jl#GeneralizedGenerated# includes facilities for lifting Julia values to the type level.
+
+The type of a Soss model is thus \jl#Model{A,B,M}#, where 
 \begin{itemize}

### 2020-01-05 paper/paper.tex
edits

@@
 
-\title{Soss}
+\title{Soss: Declarative Probabilistic Programming via Runtime Code Generation}
 %  \titlenote{This is a titlenote}
@@
 
+\begin{figure}[t!]
+    \centering
+
+    \begin{subfigure}[t]{0.4\textwidth}
+        \[
+    \begin{aligned}
+        \ell = 
+           &- 3 \log{\left(2 \right)} 
+            - 5 \log{\left(\pi \right)} 
+            - σ^{2} 
+        \\ &- 4 \log{\left(λ \right)} 
+            - 2 \log{\left(λ^{2} + 1 \right)} 
+            - \frac{σ_I^{2}}{λ^{2}} 
+            - \frac{σ_J^{2}}{λ^{2}}
+        \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
+        \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
+        \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac    {\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}    }\right)
+    \end{aligned}
+    \]
+    \caption{Direct implementation}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.4\textwidth} 
+        \[
+    \begin{aligned}
+    \ell =
+    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J - σ^{2} 
+    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
+        - I J \log{σ } 
+    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
+    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
+        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
+    \\ &- \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
+    \end{aligned}
+\]
+    \captionsetup{skip=27pt}
+    \caption{After symbolic simplification}
+    \end{subfigure}
+    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps.}
+    \label{fig:symbolic}
+\end{figure}
+
+Figure~\ref{fig:symbolic} show the simplification performed automatically by \jl#symlogpdf#. Note the reduction in the terms included in summations, which account for the bulk of evaluation cost.
+
 \subsection{Inference}
@@
 
-\todo[inline]{Restructure my writing?}
+% \todo[inline]{Restructure my writing?}
 
-Besides the purely static code generation via regular macros happening in parsing time, Soss heavily uses a mechanism of "zero-cost" runtime code generation originated by Julia's generated functions \cite{bezanson2012julia}, a.k.a staged functions in the original paper and earlier versions of Julia Programming Language.
+% Unfortunately, in current stage, there're some implementation restrictions to Julia's generated functions, and generated functions cannot generate arbitrary Julia codes,
+% where some advanced programming constructions are missing, such as generators and function-related stuffs like closures(including functions with no free variables), multiple dispatch, etc.
 
-The generated functions provide the capability of performing code generation during type inference, generating programs computed by the body of function(a.k.a, generator), once and only once for each combination of argument types.
+% To address this, we introduced the works of easing the restrictions of Julia generated functions,
+% \todo[inline]{cite GG here?}
+% which allow us to generate closures from generated functions, and make the programs sufficiently powerful.
 
-Further, Julia enables type inference and compiler optimizations equivalent to non-runtime ones in runtime for the callsites of generated functions, hence we gain runtime code generation without losing performance.
-
-To take advantage of this, Soss designs a system to encode sufficient information for generating actual codes, into the types, or objects that can be dispatched like types, which allows the use of generated functions here.
-
-Unfortunately, in current stage, there're some implementation restrictions to Julia's generated functions, and generated functions cannot generate arbitrary Julia codes,
-where some advanced programming constructions are missing, such as generators and function-related stuffs like closures(including functions with no free variables), multiple dispatch, etc.
-
-To address this, we introduced the works of easing the restrictions of Julia generated functions,
-\todo[inline]{cite GG here?}
-which allow us to generate closures from generated functions, and make the programs sufficiently powerful.
-
-Programs that can be generated by Soss are turing complete, as all constructs of Lambda Calculus can be generated.
+% Programs that can be generated by Soss are turing complete, as all constructs of Lambda Calculus can be generated.
 
@@
 %   \end{minted}
-
-\section{Extending Soss}
 
@@
 \section{Symbolic Simplification}
-
-\begin{figure}[t!]
-    \centering
-
-    \begin{subfigure}[t]{0.4\textwidth}
-        \[
-    \begin{aligned}
-        \ell = 
-           &- 3 \log{\left(2 \right)} 
-            - 5 \log{\left(\pi \right)} 
-            - σ^{2} 
-        \\ &- 4 \log{\left(λ \right)} 
-            - 2 \log{\left(λ^{2} + 1 \right)} 
-            - \frac{σ_I^{2}}{λ^{2}} 
-            - \frac{σ_J^{2}}{λ^{2}}
-        \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
-        \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
-        \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac    {\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}    }\right)
-    \end{aligned}
-    \]
-    \caption{Direct implementation}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.4\textwidth} 
-        \[
-    \begin{aligned}
-    \ell =
-    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J - σ^{2} 
-    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
-        - I J \log{σ } 
-    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
-    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
-        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
-    \\ &- \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
-    \end{aligned}
-\]
-    \captionsetup{skip=27pt}
-    \caption{After symbolic simplification}
-    \end{subfigure}
-    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps.}
-    \label{}
-\end{figure}
 

### 2020-01-05 paper/paper.tex
edits

@@
     \item[\texttt{rand}] draws a random point or sample from a given distribution.
-    \item[\texttt{particles}] draws a \emph{systematic sample}, as described in \cite{Douc2005}.
+    \item[\texttt{particles}] uses \jl#MonteCarloMeasurements.jl# to draw a \emph{systematic sample} \cite{Douc2005}.
     \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
@@
 In the above, a \emph{distribution} refers to a joint distribution, while \emph{point} refers to a named tuple in the space space of a distribution.
-
-\todo[inline]{cite MonteCarloMeasurements.jl and TransformVariables.jl}
 
@@
 
-\emph{Turing} \cite{ge2018t}
+\emph{Turing} \cite{ge2018t} is syntactically similar to Soss, but favors a more imperative style.
+Turing can represent a wider variety of models than Soss, in particular Bayesian nonparameteric models, but sacrifices some ability to reason staticically about models in order to achieve this.
+We hope to soon find ways to interface with Turing so users can benefit from both Turing's modeling flexibility and Soss's static analysis and specialized code generation.
 
@@
 
-\section{Model Syntax}
+% \section{Model Syntax}
 
-\begin{figure}[!t]
-  \centering
-  %\fbox{\rule[-.5cm]{0cm}{\linedwid} \rule[-.5cm]{4cm}{0cm}}
-  % using the backnaur package
-\begin{bnf*}
-  \bnfprod{model}{\bnfts{@model} \bnfsp \bnfpn{args} \bnfsp \bnfts{begin} \bnfsp \bnfpn{statements} \bnfsp \bnfts{end}}\\
-  \bnfprod{args}{\bnfes \bnfor \bnfpn{Symbol} \bnfor \bnfpn{Symbol} \bnfts{,} \bnfpn{args}} \\
-  \bnfprod{statements}{\bnfpn{statement} \bnfor \bnfpn{statement} \bnfsp \bnftd{Julia Line Sep} \bnfsp \bnfpn{statements} \bnfor \bnfpn{statements} \bnfsp \bnfpn{retn}} \\ 
-  \bnfprod{statement}{\bnfpn{assign} \bnfor \bnfpn{sample}} \\
-  \bnfprod{assign}{\bnfpn{Symbol} \bnfsp \bnfts{=} \bnfsp \bnfpn{Expr}} \\
-  \bnfprod{sample}{\bnfpn{Symbol} \bnfsp \bnfts{\textasciitilde}} \bnfsp \bnfpn{Measure} \\
-  \bnfprod{retn}{\bnfts{return} \bnfsp \bnfpn{Expr}} \\
-  \bnfprod{Symbol}{\bnftd{Julia Symbol}} \\ 
-  \bnfprod{Expr}{\bnftd{Julia Expr}} \\
-  \bnfprod{Measure}{\bnftd{Probability measure (see text)}}
-\end{bnf*}
-  \caption{Backus-Naur Form representation for a user-specified model.}
-  \label{fig:bnf}
-  \Description{Placeholder figure.}
-\end{figure}
-
-\section{Symbolic Simplification}
+% \begin{figure}[!t]
+%   \centering
+%   %\fbox{\rule[-.5cm]{0cm}{\linedwid} \rule[-.5cm]{4cm}{0cm}}
+%   % using the backnaur package
+% \begin{bnf*}
+%   \bnfprod{model}{\bnfts{@model} \bnfsp \bnfpn{args} \bnfsp \bnfts{begin} \bnfsp \bnfpn{statements} \bnfsp \bnfts{end}}\\
+%   \bnfprod{args}{\bnfes \bnfor \bnfpn{Symbol} \bnfor \bnfpn{Symbol} \bnfts{,} \bnfpn{args}} \\
+%   \bnfprod{statements}{\bnfpn{statement} \bnfor \bnfpn{statement} \bnfsp \bnftd{Julia Line Sep} \bnfsp \bnfpn{statements} \bnfor \bnfpn{statements} \bnfsp \bnfpn{retn}} \\ 
+%   \bnfprod{statement}{\bnfpn{assign} \bnfor \bnfpn{sample}} \\
+%   \bnfprod{assign}{\bnfpn{Symbol} \bnfsp \bnfts{=} \bnfsp \bnfpn{Expr}} \\
+%   \bnfprod{sample}{\bnfpn{Symbol} \bnfsp \bnfts{\textasciitilde}} \bnfsp \bnfpn{Measure} \\
+%   \bnfprod{retn}{\bnfts{return} \bnfsp \bnfpn{Expr}} \\
+%   \bnfprod{Symbol}{\bnftd{Julia Symbol}} \\ 
+%   \bnfprod{Expr}{\bnftd{Julia Expr}} \\
+%   \bnfprod{Measure}{\bnftd{Probability measure (see text)}}
+% \end{bnf*}
+%   \caption{Backus-Naur Form representation for a user-specified model.}
+%   \label{fig:bnf}
+%   \Description{Placeholder figure.}
+% \end{figure}
 

### 2020-01-06 paper/paper.tex
edits

@@
 \begin{acks}
-We would like to acknowledge...
+The authors would like to thank the Julia community, in particular Seth Axen, Tamas Papp, Fredrik Bagge Carlson, the Turing and Gen development teams, and the Julia founders and core developers. 
 \end{acks}
@@
 
-\begin{figure}[t]
-\centering
-\begin{subfigure}[t]{0.48\textwidth}
-\begin{minted}{julia1}
-julia> Soss.sourceRand()(m)
-quote
-    σ = rand(HalfNormal())
-    λ = rand(HalfCauchy())
-    σJ = rand(HalfNormal(λ))
-    σI = rand(HalfNormal(λ))
-    β = rand(iid(J, Normal(0, σJ)))
-    α = rand(iid(I, Normal(0, σI)))
-    yhat = α .+ β'
-    y = rand(For(I, J) do i, j
-                Normal(yhat[i, j], σ)
-            end)
-    (I=I, J=J, yhat=yhat, λ=λ, σI=σI, 
-    σJ=σJ, α=α, β=β, σ=σ, y=y) 
-end
-\end{minted}
-\caption{\footnotesize{\texttt{rand}}}
-\end{subfigure}
-\begin{subfigure}[t]{0.48\textwidth}
-\begin{minted}{julia1}
-julia> Soss.sourceLogpdf()(m)
-quote
-    _ℓ = 0.0
-    _ℓ += logpdf(HalfNormal(), σ)
-    _ℓ += logpdf(HalfCauchy(), λ)
-    _ℓ += logpdf(HalfNormal(λ), σJ)
-    _ℓ += logpdf(HalfNormal(λ), σI)
-    _ℓ += logpdf(Normal(0, σJ) |> iid(J), β)
-    _ℓ += logpdf(Normal(0, σI) |> iid(I), α)
-    yhat = α .+ β'
-    _ℓ += logpdf(For(I, J) do i, j
-                Normal(yhat[i, j], σ)
-            end, y)
-    return _ℓ
-end
-\end{minted}
-\caption{\footnotesize{\texttt{logpdf}}}
-\end{subfigure}
-\caption{Generated code when calling \footnotesize{\texttt{rand(m)}} and \footnotesize{\texttt{logpdf(m,x)}}.} 
-\label{} 
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \begin{subfigure}[t]{0.48\textwidth}
+% \begin{minted}{julia1}
+% julia> Soss.sourceRand()(m)
+% quote
+%     σ = rand(HalfNormal())
+%     λ = rand(HalfCauchy())
+%     σJ = rand(HalfNormal(λ))
+%     σI = rand(HalfNormal(λ))
+%     β = rand(iid(J, Normal(0, σJ)))
+%     α = rand(iid(I, Normal(0, σI)))
+%     yhat = α .+ β'
+%     y = rand(For(I, J) do i, j
+%                 Normal(yhat[i, j], σ)
+%             end)
+%     (I=I, J=J, yhat=yhat, λ=λ, σI=σI, 
+%     σJ=σJ, α=α, β=β, σ=σ, y=y) 
+% end
+% \end{minted}
+% \caption{\footnotesize{\texttt{rand}}}
+% \end{subfigure}
+% \begin{subfigure}[t]{0.48\textwidth}
+% \begin{minted}{julia1}
+% julia> Soss.sourceLogpdf()(m)
+% quote
+%     _ℓ = 0.0
+%     _ℓ += logpdf(HalfNormal(), σ)
+%     _ℓ += logpdf(HalfCauchy(), λ)
+%     _ℓ += logpdf(HalfNormal(λ), σJ)
+%     _ℓ += logpdf(HalfNormal(λ), σI)
+%     _ℓ += logpdf(Normal(0, σJ) |> iid(J), β)
+%     _ℓ += logpdf(Normal(0, σI) |> iid(I), α)
+%     yhat = α .+ β'
+%     _ℓ += logpdf(For(I, J) do i, j
+%                 Normal(yhat[i, j], σ)
+%             end, y)
+%     return _ℓ
+% end
+% \end{minted}
+% \caption{\footnotesize{\texttt{logpdf}}}
+% \end{subfigure}
+% \caption{Generated code when calling \footnotesize{\texttt{rand(m)}} and \footnotesize{\texttt{logpdf(m,x)}}.} 
+% \label{} 
+% \end{figure}
 
-\section{Extended example }
-\todo[inline]{Break this up! This is just a staging area}
+% \section{Extended example }
 
-\begin{minted}{julia1}
-julia> using Revise, Soss, Random
+% \begin{minted}{julia1}
+% julia> using Revise, Soss, Random
 
-julia> Random.seed!(1);
+% julia> Random.seed!(1);
 
-julia> m = @model I,J begin
-           λ ~ HalfCauchy()
-           σI ~ HalfNormal(λ)
-           σJ ~ HalfNormal(λ)
-           α ~ Normal(0, σI) |> iid(I)
-           β ~ Normal(0, σJ) |> iid(J)
-           yhat = α .+ β'
-           σ ~ HalfNormal()
-           y ~ For(I,J) do i,j 
-                   Normal(yhat[i,j], σ)
-               end
-       end;
+% julia> m = @model I,J begin
+%            λ ~ HalfCauchy()
+%            σI ~ HalfNormal(λ)
+%            σJ ~ HalfNormal(λ)
+%            α ~ Normal(0, σI) |> iid(I)
+%            β ~ Normal(0, σJ) |> iid(J)
+%            yhat = α .+ β'
+%            σ ~ HalfNormal()
+%            y ~ For(I,J) do i,j 
+%                    Normal(yhat[i,j], σ)
+%                end
+%        end;
 
-julia> truth = rand(m(I=2,J=3)); pairs(truth);
+% julia> truth = rand(m(I=2,J=3)); pairs(truth);
 
-julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));
+% julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));
 
-julia> postpred = predict(m(I=2,J=3), post) |> particles;
+% julia> postpred = predict(m(I=2,J=3), post) |> particles;
 
-julia> postpred.yhat
-2×3 Array{Particles{Float64,1000},2}:
- 0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
- 0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37
+% julia> postpred.yhat
+% 2×3 Array{Particles{Float64,1000},2}:
+%  0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
+%  0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37
 
-julia> postpred.y
-2×3 Array{Particles{Float64,1000},2}:
- 0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
- 0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61
+% julia> postpred.y
+% 2×3 Array{Particles{Float64,1000},2}:
+%  0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
+%  0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61
 
-julia> m.args
-2-element Array{Symbol,1}:
- :I
- :J
+% julia> m.args
+% 2-element Array{Symbol,1}:
+%  :I
+%  :J
 
-julia> pairs(m.vals)
-pairs(::NamedTuple) with 1 entry:
-  :yhat => :(α .+ β')
+% julia> pairs(m.vals)
+% pairs(::NamedTuple) with 1 entry:
+%   :yhat => :(α .+ β')
 
-julia> pairs(m.dists)
-pairs(::NamedTuple) with 7 entries:
-  :λ  => :(HalfCauchy())
-  :σI => :(HalfNormal(λ))
-  :σJ => :(HalfNormal(λ))
-  :α  => :(Normal(0, σI) |> iid(I))
-  :β  => :(Normal(0, σJ) |> iid(J))
-  :σ  => :(HalfNormal())
-  :y  => :(For(I, J) do i, j…
+% julia> pairs(m.dists)
+% pairs(::NamedTuple) with 7 entries:
+%   :λ  => :(HalfCauchy())
+%   :σI => :(HalfNormal(λ))
+%   :σJ => :(HalfNormal(λ))
+%   :α  => :(Normal(0, σI) |> iid(I))
+%   :β  => :(Normal(0, σJ) |> iid(J))
+%   :σ  => :(HalfNormal())
+%   :y  => :(For(I, J) do i, j…
 
-\end{minted}
+% \end{minted}
 

### 2020-01-10 paper/paper.tex
Add Taine's email
edits

@@
 }
-% \email{apfeffer@cra.com}
+\email{thaut@logic.cs.tsukuba.ac.jp}
 %\renewcommand\shortauthors{Mage, M. et al}
@@
     \item[\texttt{rand}] draws a random point or sample from a given distribution.
-    \item[\texttt{particles}] uses \jl#MonteCarloMeasurements.jl# to draw a \emph{systematic sample} \cite{Douc2005}.
+    \item[\texttt{particles}] draws a \emph{systematic sample} using \jl#MonteCarloMeasurements.jl# \cite{8ff6a743-0ad6-4d98-bbb3-5d549c698bc1}.
     \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
@@
 
-\emph{Turing} \cite{ge2018t} is syntactically similar to Soss, but favors a more imperative style.
-Turing can represent a wider variety of models than Soss, in particular Bayesian nonparameteric models, but sacrifices some ability to reason staticically about models in order to achieve this.
+\emph{Turing} \cite{ge2018t} is syntactically similar to Soss, but tends to favor a more imperative style.
+Turing calls the Julia compiler at a much earlier stage than Soss.
+This allows representation of a wider variety of models, at the cost of some flexibility of model transformations.
 We hope to soon find ways to interface with Turing so users can benefit from both Turing's modeling flexibility and Soss's static analysis and specialized code generation.

### 2021-09-21 paper/paper.tex
final version

@@
                format=acmsmall, %
-               review=true, %
+               review=false, %
                screen=true, %
@@
   %\postcode{02139}
-  %\country{USA}
+  \country{USA}
 }
@@
   %\postcode{02138}
-  %\country{USA}
+  \country{Japan}
 }
@@
     \item[\texttt{rand}] draws a random point or sample from a given distribution.
-    \item[\texttt{particles}] draws a \emph{systematic sample} using \jl#MonteCarloMeasurements.jl# \cite{8ff6a743-0ad6-4d98-bbb3-5d549c698bc1}.
+    \item[\texttt{particles}] draws a \emph{systematic sample} using \jl#MonteCarloMeasurements.jl# \cite{bagge2020MCM.jl}.
     \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
